---
title: "Ridge & Lasso using Glmnet"
author: "John Zhuang"
date: "2024-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Data Preparation

```{r}
library(glmnet)
```

```{r}
data.horizontal = readRDS("data/derived/data_horizontal_joined.rds")
```

## A. Check for missingness of columns
```{r}
missing = colSums(is.na(data.horizontal))
missing[which(missing > 0)]
```

These columns have missing data and will thus be removed from the data.

```{r}
data.glmnet = data.horizontal[, which(missing == 0)]
```

# 1. Ridge, Lasso and Elastic Net using Glmnet

```{r}
## Perform cross validation using glmnet in order to find the best lambda
## for shrinkage, selection or both.

vs.cv.glm = function(alpha, measure, fold = 10) {
                    # Column 1000 is vital_status, the disease outcome
  model = cv.glmnet(x = data.matrix(data.glmnet[, !names(data.glmnet) %in% c("vital_status", "event_time")]), 
                    y = data.matrix(data.glmnet$vital_status),
                    alpha = alpha, # 0 for Ridge, 1 for Lasso, 0 < alpha < 1 for Elastic Net
                    family = "binomial", # Outcome is Binomial
                    type.measure = measure,
                    nfolds = fold,
                    standardize = T)
}
```

## A. 10-fold Glmnet

```{r}
set.seed(42)
mse.0_10 = vs.cv.glm(0, measure = "mse")
mse.0.5_10 = vs.cv.glm(0.5, measure = "mse")
mse.1_10 = vs.cv.glm(1, measure = "mse")

dev.0_10 = vs.cv.glm(0, measure = "deviance")
dev.0.5_10 = vs.cv.glm(0.5, measure = "deviance")
dev.1_10 = vs.cv.glm(1, measure = "deviance")

#Warning: Too few (< 10) observations per fold for type.measure='auc'

auc.0_10 = vs.cv.glm(0, measure = "auc")
auc.0.5_10 = vs.cv.glm(0.5, measure = "auc")
auc.1_10 = vs.cv.glm(1, measure = "auc")
```

```{r}
out.ridge_10 = list("mse" = mse.0_10,
                "dev" = dev.0_10)#,
                #"auc" = auc.0_10)

out.elasticnet_10 = list("mse" = mse.0.5_10,
                         "dev" = dev.0.5_10)#,
                         #"auc" = auc.0.5_10)

out.lasso_10 = list("mse" = mse.1_10,
                "dev" = dev.1_10)#,
                #"auc" = auc.1_10)

out.glmnet_10 = list("ridge" = out.ridge_10,
                 "elasticnet_0.5" = out.elasticnet_10,
                 "lasso" = out.lasso_10)
```

```{r}
saveRDS(out.glmnet_10, file = "data/out_models/results_glmnet_10folds.rds")
```

## B. 5-fold Glmnet

```{r}
set.seed(42)
mse.0_5 = vs.cv.glm(0, measure = "mse", fold = 5)
mse.0.5_5 = vs.cv.glm(0.5, measure = "mse", fold = 5)
mse.1_5 = vs.cv.glm(1, measure = "mse", fold = 5)

dev.0_5 = vs.cv.glm(0, measure = "deviance", fold = 5)
dev.0.5_5 = vs.cv.glm(0.5, measure = "deviance", fold = 5)
dev.1_5 = vs.cv.glm(1, measure = "deviance", fold = 5)

#Warning: Too few (< 10) observations per fold for type.measure='auc'

auc.0_5 = vs.cv.glm(0, measure = "auc", fold = 5)
auc.0.5_5 = vs.cv.glm(0.5, measure = "auc", fold = 5)
auc.1_5 = vs.cv.glm(1, measure = "auc", fold = 5)
```

```{r}
out.ridge_5 = list("mse" = mse.0_5,
                "dev" = dev.0_5)#,
                #"auc" = auc.0_5)

out.elasticnet_5 = list("mse" = mse.0.5_5,
                         "dev" = dev.0.5_5)#,
                         #"auc" = auc.0.5_5)

out.lasso_5 = list("mse" = mse.1_5,
                "dev" = dev.1_5)#,
                #"auc" = auc.1_5)

out.glmnet_5 = list("ridge" = out.ridge_5,
                 "elasticnet_0.5" = out.elasticnet_5,
                 "lasso" = out.lasso_5)
```

```{r}
saveRDS(out.glmnet_5, file = "data/out_models/results_glmnet_5folds.rds")
```

## C. 3-fold Glmnet

```{r}
set.seed(42)
mse.0_3 = vs.cv.glm(0, measure = "mse", fold = 3)
mse.0.5_3 = vs.cv.glm(0.5, measure = "mse", fold = 3)
mse.1_3 = vs.cv.glm(1, measure = "mse", fold = 3)

dev.0_3 = vs.cv.glm(0, measure = "deviance", fold = 3)
dev.0.5_3 = vs.cv.glm(0.5, measure = "deviance", fold = 3)
dev.1_3 = vs.cv.glm(1, measure = "deviance", fold = 3)


auc.0_3 = vs.cv.glm(0, measure = "auc", fold = 3)
auc.0.5_3 = vs.cv.glm(0.5, measure = "auc", fold = 3)
auc.1_3 = vs.cv.glm(1, measure = "auc", fold = 3)
```

```{r}
out.ridge_3 = list("mse" = mse.0_3,
                "dev" = dev.0_3,
                "auc" = auc.0_3)

out.elasticnet_3 = list("mse" = mse.0.5_3,
                         "dev" = dev.0.5_3,
                         "auc" = auc.0.5_3)

out.lasso_3 = list("mse" = mse.1_3,
                "dev" = dev.1_3,
                "auc" = auc.1_3)

out.glmnet_3 = list("ridge" = out.ridge_3,
                 "elasticnet_0.5" = out.elasticnet_3,
                 "lasso" = out.lasso_3)
```

```{r}
saveRDS(out.glmnet_3, file = "data/out_models/results_glmnet_3folds.rds")
```

# 2. Results Presentation

```{r}
print.coef = function(cv.glm, coef.num = T) {
  coef.cv.glm = coef(cv.glm, s="lambda.min") 
  out = coef.cv.glm[coef.cv.glm[,1] != 0,]
  if (!coef.num) {
    out = names(out)
  }
  out
}
```

## A. 10 folds

### Ridge

```{r}
plot(mse.0_10)
title("10-fold CV, alpha = 0", line = 2.5)
plot(dev.0_10)
title("10-fold CV, alpha = 0", line = 2.5)
plot(auc.0_10) # Actually Deviance
title("10-fold CV, alpha = 0", line = 2.5)
```

### Elastic Net (Alpha =  0.5)

```{r}
plot(mse.0.5_10)
title("10-fold CV, alpha = 0.5", line = 2.5)
plot(dev.0.5_10)
title("10-fold CV, alpha = 0.5", line = 2.5)
plot(auc.0.5_10) # Actually Deviance
title("10-fold CV, alpha = 0.5", line = 2.5)

print.coef(mse.0.5_10)
print.coef(dev.0.5_10)
print.coef(auc.0.5_10)
```

### Lasso

```{r}
plot(mse.1_10)
title("10-fold CV, alpha = 1", line = 2.5)
plot(dev.1_10)
title("10-fold CV, alpha = 1", line = 2.5)
plot(auc.1_10) # Actually Deviance
title("10-fold CV, alpha = 1", line = 2.5)

print.coef(mse.1_10)
print.coef(dev.1_10)
print.coef(auc.1_10)
```

## B. 5 folds

### Ridge

```{r}
plot(mse.0_5)
title("5-fold CV, alpha = 0", line = 2.5)
plot(dev.0_5)
title("5-fold CV, alpha = 0", line = 2.5)
plot(auc.0_5) # Actually Deviance
title("5-fold CV, alpha = 0", line = 2.5)
```

### Elastic Net (Alpha =  0.5)

```{r}
plot(mse.0.5_5)
title("5-fold CV, alpha = 0.5", line = 2.5)
plot(dev.0.5_5)
title("5-fold CV, alpha = 0.5", line = 2.5)
plot(auc.0.5_5)
title("5-fold CV, alpha = 0.5", line = 2.5)

print.coef(mse.0.5_5)
print.coef(dev.0.5_5)
print.coef(auc.0.5_5)
```

### Lasso

```{r}
plot(mse.1_5)
title("5-fold CV, alpha = 1", line = 2.5)
plot(dev.1_5)
title("5-fold CV, alpha = 1", line = 2.5)
plot(auc.1_5) # Actually Deviance
title("5-fold CV, alpha = 1", line = 2.5)

print.coef(mse.1_5)
print.coef(dev.1_5)
print.coef(auc.1_5)
```

## C. 3 folds

### Ridge

```{r}
plot(mse.0_3)
title("3-fold CV, alpha = 0", line = 2.5)
plot(dev.0_3)
title("3-fold CV, alpha = 0", line = 2.5)
plot(auc.0_3)
title("3-fold CV, alpha = 0", line = 2.5)
```

### Elastic Net (Alpha =  0.5)

```{r}
plot(mse.0.5_3)
title("3-fold CV, alpha = 0.5", line = 2.5)
plot(dev.0.5_3)
title("3-fold CV, alpha = 0.5", line = 2.5)
plot(auc.0.5_3)
title("3-fold CV, alpha = 0.5", line = 2.5)

print.coef(mse.0.5_3)
print.coef(dev.0.5_3)
print.coef(auc.0.5_3)
```

### Lasso

```{r}
plot(mse.1_3)
title("3-fold CV, alpha = 1", line = 2.5)
plot(dev.1_3)
title("3-fold CV, alpha = 1", line = 2.5)
plot(auc.1_3)
title("3-fold CV, alpha = 1", line = 2.5)

print.coef(mse.1_3)
print.coef(dev.1_3)
print.coef(auc.1_3)
```

# 3. Further Alpha attempts

We will explore alphas < 0.5 for 10 and 5 folds Glmnet, as they return 0 selected features for Elastic net (alpha = 0.5) and Lasso.

It was found that alpha = 0.2 returns enough selected features for 10 and 5 folds Glmnet.

## A. 10 folds

```{r}
set.seed(42)
plot(vs.cv.glm(0.1, measure = "mse"))
title("10-fold CV, alpha = 0.1", line = 2.5)
plot(vs.cv.glm(0.2, measure = "mse"))
title("10-fold CV, alpha = 0.2", line = 2.5)
plot(vs.cv.glm(0.3, measure = "mse"))
title("10-fold CV, alpha = 0.3", line = 2.5)
plot(vs.cv.glm(0.4, measure = "mse"))
title("10-fold CV, alpha = 0.4", line = 2.5)
```
```{r}
set.seed(42)
plot(vs.cv.glm(0.1, measure = "deviance"))
title("10-fold CV, alpha = 0.1", line = 2.5)
plot(vs.cv.glm(0.2, measure = "deviance"))
title("10-fold CV, alpha = 0.2", line = 2.5)
plot(vs.cv.glm(0.3, measure = "deviance"))
title("10-fold CV, alpha = 0.3", line = 2.5)
plot(vs.cv.glm(0.4, measure = "deviance"))
title("10-fold CV, alpha = 0.4", line = 2.5)
```


```{r}
set.seed(42)
mse.0.2_10 = vs.cv.glm(0.2, measure = "mse")
dev.0.2_10 = vs.cv.glm(0.2, measure = "deviance")
auc.0.2_10 = vs.cv.glm(0.2, measure = "auc")

plot(mse.0.2_10)
title("10-fold CV, alpha = 0.2", line = 2.5)
plot(dev.0.2_10)
title("10-fold CV, alpha = 0.2", line = 2.5)
plot(auc.0.2_10)
title("10-fold CV, alpha = 0.2", line = 2.5)
print.coef(mse.0.2_10)
print.coef(dev.0.2_10)
print.coef(auc.0.2_10)
```

```{r}
out.glmnet.0.2_10 = list("mse" = mse.0.2_10,
                      "dev" = dev.0.2_10)

saveRDS(out.glmnet.0.2_10, file = "data/out_models/results_glmnet_0.2_10folds.rds")
```

## B. 5 folds

```{r}
set.seed(42)
plot(vs.cv.glm(0.1, measure = "mse", fold = 5))
title("5-fold CV, alpha = 0.1", line = 2.5)
plot(vs.cv.glm(0.2, measure = "mse", fold = 5))
title("5-fold CV, alpha = 0.2", line = 2.5)
plot(vs.cv.glm(0.3, measure = "mse", fold = 5))
title("5-fold CV, alpha = 0.3", line = 2.5)
plot(vs.cv.glm(0.4, measure = "mse", fold = 5))
title("5-fold CV, alpha = 0.4", line = 2.5)
```
```{r}
set.seed(42)
plot(vs.cv.glm(0.1, measure = "deviance", fold = 5))
title("5-fold CV, alpha = 0.1", line = 2.5)
plot(vs.cv.glm(0.2, measure = "deviance", fold = 5))
title("5-fold CV, alpha = 0.2", line = 2.5)
plot(vs.cv.glm(0.3, measure = "deviance", fold = 5))
title("5-fold CV, alpha = 0.3", line = 2.5)
plot(vs.cv.glm(0.4, measure = "deviance", fold = 5))
title("5-fold CV, alpha = 0.4", line = 2.5)
```

```{r}
set.seed(42)
mse.0.2_5 = vs.cv.glm(0.2, measure = "mse", fold = 5)
dev.0.2_5 = vs.cv.glm(0.2, measure = "deviance", fold = 5)
auc.0.2_5 = vs.cv.glm(0.2, measure = "auc", fold = 5)

plot(mse.0.2_5)
title("5-fold CV, alpha = 0.2", line = 2.5)
plot(dev.0.2_5)
title("5-fold CV, alpha = 0.2", line = 2.5)
plot(auc.0.2_5)
title("5-fold CV, alpha = 0.2", line = 2.5)
print.coef(mse.0.2_5)
print.coef(dev.0.2_5)
print.coef(auc.0.2_5)
```


```{r}
out.glmnet.0.2_5 = list("mse" = mse.0.2_5,
                      "dev" = dev.0.2_5)

saveRDS(out.glmnet.0.2_5, file = "data/out_models/results_glmnet_0.2_5folds.rds")
```

Save all coefficients as a separate RDS file for easy access:

```{r}
out.coef.0.2_5 = list("mse" = print.coef(mse.0.2_5),
                      "dev" = print.coef(dev.0.2_5))

out.coef.0.5_3 = list("mse" = print.coef(mse.0.5_3),
                      "dev" = print.coef(dev.0.5_3),
                      "auc" = print.coef(auc.0.5_3))

out.coef.1_3 = list("mse" = print.coef(mse.1_3),
                    "dev" = print.coef(dev.1_3),
                    "auc" = print.coef(auc.1_3))

out.coef.0.5_5 = list("mse" = print.coef(mse.0.5_5),
                      "dev" = print.coef(dev.0.5_5))

out.coef.1_5 = list("mse" = print.coef(mse.1_5),
                    "dev" = print.coef(dev.1_5))

out.coef.0.2_10 = list("mse" = print.coef(mse.0.2_10),
                      "dev" = print.coef(dev.0.2_10))

out.coef.0.5_10 = list("mse" = print.coef(mse.0.5_10),
                      "dev" = print.coef(dev.0.5_10))

out.coef.1_10 = list("mse" = print.coef(mse.1_10),
                    "dev" = print.coef(dev.1_10))

out.coef.3 = list("elasticnet_0.5" = out.coef.0.5_3,
                  "lasso" = out.coef.1_3)

out.coef.5 = list("elasticnet_0.2" = out.coef.0.2_5,
                  "elasticnet_0.5" = out.coef.0.5_5,
                  "lasso" = out.coef.1_5)

out.coef.10 = list("elasticnet_0.2" = out.coef.0.2_10,
                   "elasticnet_0.5" = out.coef.0.5_10,
                  "lasso" = out.coef.1_10)

out.coef = list("fold3" = out.coef.3,
                "fold5" = out.coef.5,
                "fold10" = out.coef.10)

saveRDS(out.coef, file = "data/sel_features/results_glmnet_coef.rds")
```

