---
title: "Survival Analysis (Cox Model) in Glmnet"
author: "John Zhuang"
date: "2024-06-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Data Preparation

```{r}
library(glmnet)
```

```{r}
data.horizontal = readRDS("data/derived/data_horizontal_joined.rds")
```

## A. Make survival outcome matrix for Survival analysis

```{r}
survival.Y = data.horizontal[, names(data.horizontal) %in% c("vital_status", "event_time")]

names(survival.Y) = c("status", "time") # Rename columns for glmnet compatibility
survival.Y$status = as.numeric(survival.Y$status) - 1 #Revert to numeric form for glmnet compatibility
survival.Y = as.matrix(survival.Y)
```

## B. Make survival independent variable matrix.

Age is the only metadata variable included because glmnet cox regression requires a matrix as input.

```{r}
survival.X = as.matrix(data.horizontal[,1:997])
```



# 1. Cox model survival analysis using Glmnet

```{r}
## Perform cross validation using glmnet in order to find the best lambda
## for shrinkage, selection or both.

cox.cv.glm = function(alpha, measure = NA, fold = 10) {
                    # Column 1000 is vital_status, the disease outcome
  model = cv.glmnet(x = survival.X, 
                    y = survival.Y,
                    alpha = alpha, # 0 for Ridge, 1 for Lasso, 0 < alpha < 1 for Elastic Net
                    family = "cox", # Survival Analysis
                    type.measure = measure, # "deviance" for partial-likelihood, "c" for Harrell C index
                    nfolds = fold,
                    standardize = T)
}
```

## 10 folds

```{r}
set.seed(42)
dev.0_10 = cox.cv.glm(0, "deviance")
dev.0.5_10 = cox.cv.glm(0.5, "deviance")
dev.1_10 = cox.cv.glm(1, "deviance")
harrell.0_10 = cox.cv.glm(0, "C")
harrell.0.5_10 = cox.cv.glm(0.5, "C")
harrell.1_10 = cox.cv.glm(1, "C")
```

```{r}
out.ridge_10 = list("dev" = dev.0_10,
                "c" = harrell.0_10)

out.elasticnet_10 = list("dev" = dev.0.5_10,
                "c" = harrell.0.5_10)

out.lasso_10 = list("dev" = dev.1_10,
                "c" = harrell.1_10)

out.cox.glmnet_10 = list("ridge" = out.ridge_10,
                 "elasticnet_0.5" = out.elasticnet_10,
                 "lasso" = out.lasso_10)
```

```{r}
saveRDS(out.cox.glmnet_10, file = "data/out_models/results_glmnet_cox_10folds.rds")
```

## 5 folds

```{r}
set.seed(42)
dev.0_5 = cox.cv.glm(0, "deviance", fold = 5)
dev.0.5_5 = cox.cv.glm(0.5, "deviance", fold = 5)
dev.1_5 = cox.cv.glm(1, "deviance", fold = 5)
harrell.0_5 = cox.cv.glm(0, "C", fold = 5)
harrell.0.5_5 = cox.cv.glm(0.5, "C", fold = 5)
harrell.1_5 = cox.cv.glm(1, "C", fold = 5)
```

```{r}
out.ridge_5 = list("dev" = dev.0_5,
                "c" = harrell.0_5)

out.elasticnet_5 = list("dev" = dev.0.5_5,
                "c" = harrell.0.5_5)

out.lasso_5 = list("dev" = dev.1_5,
                "c" = harrell.1_5)

out.cox.glmnet_5 = list("ridge" = out.ridge_5,
                 "elasticnet_0.5" = out.elasticnet_5,
                 "lasso" = out.lasso_5)
```

```{r}
saveRDS(out.cox.glmnet_5, file = "data/out_models/results_glmnet_cox_5folds.rds")
```

## 3 folds

```{r}
set.seed(42)
dev.0_3 = cox.cv.glm(0, "deviance", fold = 3)
dev.0.5_3 = cox.cv.glm(0.5, "deviance", fold = 3)
dev.1_3 = cox.cv.glm(1, "deviance", fold = 3)
harrell.0_3 = cox.cv.glm(0, "C", fold = 3)
harrell.0.5_3 = cox.cv.glm(0.5, "C", fold = 3)
harrell.1_3 = cox.cv.glm(1, "C", fold = 3)
```

```{r}
out.ridge_3 = list("dev" = dev.0_3,
                "c" = harrell.0_3)

out.elasticnet_3 = list("dev" = dev.0.5_3,
                "c" = harrell.0.5_3)

out.lasso_3 = list("dev" = dev.1_3,
                "c" = harrell.1_3)

out.cox.glmnet_3 = list("ridge" = out.ridge_3,
                 "elasticnet_0.5" = out.elasticnet_3,
                 "lasso" = out.lasso_3)
```

```{r}
saveRDS(out.cox.glmnet_3, file = "data/out_models/results_glmnet_cox_3folds.rds")
```

# 2. Results Presentation

```{r}
print.coef = function(cv.glm, coef.num = T) {
  coef.cv.glm = coef(cv.glm, s="lambda.min") 
  out = coef.cv.glm[coef.cv.glm[,1] != 0,]
  if (!coef.num) {
    out = names(out)
  }
  out
}
```

## A. 10 folds

### Ridge

```{r}
plot(dev.0_10)
title("10-fold CV, alpha = 0", line = 2.5)
plot(harrell.0_10)
title("10-fold CV, alpha = 0", line = 2.5)
```


### Elastic Net (Alpha = 0.5)

```{r}
plot(dev.0.5_10)
title("10-fold CV, alpha = 0.5", line = 2.5)
plot(harrell.0.5_10)
title("10-fold CV, alpha = 0.5", line = 2.5)

print.coef(dev.0.5_10)
print.coef(harrell.0.5_10)
```

## Lasso

```{r}
plot(dev.1_10)
title("10-fold CV, alpha = 1", line = 2.5)
plot(harrell.1_10)
title("10-fold CV, alpha = 1", line = 2.5)

print.coef(dev.1_10)
print.coef(harrell.1_10)
```

## B. 5 folds

### Ridge

```{r}
plot(dev.0_5)
title("5-fold CV, alpha = 0", line = 2.5)
plot(harrell.0_5)
title("5-fold CV, alpha = 0", line = 2.5)
```


### Elastic Net (Alpha = 0.5)

```{r}
plot(dev.0.5_5)
title("5-fold CV, alpha = 0.5", line = 2.5)
plot(harrell.0.5_5)
title("5-fold CV, alpha = 0.5", line = 2.5)

print.coef(dev.0.5_5)
print.coef(harrell.0.5_5)
```

## Lasso

```{r}
plot(dev.1_5)
title("5-fold CV, alpha = 1", line = 2.5)
plot(harrell.1_5)
title("5-fold CV, alpha = 1", line = 2.5)

print.coef(dev.1_5)
print.coef(harrell.1_5)
```

## C. 3 folds

### Ridge

```{r}
plot(dev.0_3)
title("3-fold CV, alpha = 0", line = 2.5)
plot(harrell.0_3)
title("3-fold CV, alpha = 0", line = 2.5)
```


### Elastic Net (Alpha = 0.5)

```{r}
plot(dev.0.5_3)
title("3-fold CV, alpha = 0.5", line = 2.5)
plot(harrell.0.5_3)
title("3-fold CV, alpha = 0.5", line = 2.5)

print.coef(dev.0.5_3)
print.coef(harrell.0.5_3)
```

## Lasso

```{r}
plot(dev.1_3)
title("3-fold CV, alpha = 1", line = 2.5)
plot(harrell.1_3)
title("3-fold CV, alpha = 1", line = 2.5)

print.coef(dev.1_3)
print.coef(harrell.1_3)
```

Save all coefficients as a separate RDS file for easy access:

```{r}

out.coef.0.5_3 = list("dev" = print.coef(dev.0.5_3),
                      "c" = print.coef(harrell.0.5_3))

out.coef.1_3 = list("dev" = print.coef(dev.1_3),
                      "c" = print.coef(harrell.1_3))

out.coef.0.5_5 = list("dev" = print.coef(dev.0.5_5),
                      "c" = print.coef(harrell.0.5_5))

out.coef.1_5 = list("dev" = print.coef(dev.1_5),
                      "c" = print.coef(harrell.1_5))

out.coef.0.5_10 = list("dev" = print.coef(dev.0.5_10),
                      "c" = print.coef(harrell.0.5_10))

out.coef.1_10 = list("dev" = print.coef(dev.1_10),
                      "c" = print.coef(harrell.1_10))

out.coef.3 = list("elasticnet_0.5" = out.coef.0.5_3,
                  "lasso" = out.coef.1_3)

out.coef.5 = list("elasticnet_0.5" = out.coef.0.5_5,
                  "lasso" = out.coef.1_5)

out.coef.10 = list("elasticnet_0.5" = out.coef.0.5_10,
                  "lasso" = out.coef.1_10)

out.coef = list("fold3" = out.coef.3,
                "fold5" = out.coef.5,
                "fold10" = out.coef.10)

saveRDS(out.coef, file = "data/sel_features/results_glmnet_cox_coef.rds")
```

# 3. 1 standard deviation Lambda

The largest value of lambda such that the error is within 1 standard error of the minimum.

The result might be less sensitive to outsiders.

```{r}
print.coef.1se = function(cv.glm, coef.num = T) {
  coef.cv.glm = coef(cv.glm, s="lambda.1se") 
  out = coef.cv.glm[coef.cv.glm[,1] != 0,]
  if (!coef.num) {
    out = names(out)
  }
  out
}
```

## A. 10 folds


### Elastic Net (Alpha = 0.5)

```{r}
print.coef.1se(dev.0.5_10)
print.coef.1se(harrell.0.5_10)
```

## Lasso

```{r}
print.coef.1se(dev.1_10)
print.coef.1se(harrell.1_10)
```

## B. 5 folds


### Elastic Net (Alpha = 0.5)

```{r}
print.coef.1se(dev.0.5_5)
print.coef.1se(harrell.0.5_5)
```

## Lasso

```{r}

print.coef.1se(dev.1_5)
print.coef.1se(harrell.1_5)
```

## C. 3 folds

### Elastic Net (Alpha = 0.5)

```{r}

print.coef.1se(dev.0.5_3)
print.coef.1se(harrell.0.5_3)
```

## Lasso

```{r}

print.coef.1se(dev.1_3)
print.coef.1se(harrell.1_3)
```

Save all coefficients as a separate RDS file for easy access:

```{r}

out.coef.0.5_3.1se = list("dev" = print.coef.1se(dev.0.5_3),
                      "c" = print.coef.1se(harrell.0.5_3))

out.coef.1_3.1se = list("dev" = print.coef.1se(dev.1_3),
                      "c" = print.coef.1se(harrell.1_3))

out.coef.0.5_5.1se = list("dev" = print.coef.1se(dev.0.5_5),
                      "c" = print.coef.1se(harrell.0.5_5))

out.coef.1_5.1se = list("dev" = print.coef.1se(dev.1_5),
                      "c" = print.coef.1se(harrell.1_5))

out.coef.0.5_10.1se = list("dev" = print.coef.1se(dev.0.5_10),
                      "c" = print.coef.1se(harrell.0.5_10))

out.coef.1_10.1se = list("dev" = print.coef.1se(dev.1_10),
                      "c" = print.coef.1se(harrell.1_10))

out.coef.3.1se = list("elasticnet_0.5" = out.coef.0.5_3.1se,
                  "lasso" = out.coef.1_3.1se)

out.coef.5.1se = list("elasticnet_0.5" = out.coef.0.5_5.1se,
                  "lasso" = out.coef.1_5.1se)

out.coef.10.1se = list("elasticnet_0.5" = out.coef.0.5_10.1se,
                  "lasso" = out.coef.1_10.1se)

out.coef.1se = list("fold3" = out.coef.3.1se,
                "fold5" = out.coef.5.1se,
                "fold10" = out.coef.10.1se)

saveRDS(out.coef.1se, file = "data/sel_features/results_glmnet_cox_coef_1se.rds")
```

