---
title: "FABIA"
author: "John Zhuang"
date: "2024-05-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(fabia)
library(ggplot2)
library(ggrepel)
```

```{r}
data.vertical = readRDS("data/derived/data_vertical_numeric_only.rds")
```

# 0. Preparation

## a) Check again to reassure if colnames are identical before binding
```{r}
#
identical(colnames(data.vertical$RPPAArray), colnames(data.vertical$RNASeq2GeneNorm))
identical(colnames(data.vertical$RNASeq2GeneNorm), colnames(data.vertical$miRNASeqGene))
```
## b) Merge data into 1 matrix
```{r}
data.vert.joined = do.call("rbind", list(data.vertical$RPPAArray, data.vertical$RNASeq2GeneNorm, data.vertical$miRNASeqGene))
```
```{r}
data.fabia = as.matrix(data.vert.joined)
```


# 1. FABIA

## 1 Bicluster

```{r}
set.seed(42)
fabia_1 = fabia(data.fabia, p = 1)
```
```{r}
fabia_1
summary(fabia_1)
#plot(fabia_1) # Minimum 2 Dimensions needed
# extractPlot(fabia_1) # Cannot plot loadings & samples
```

## 2 Biclusters
```{r}
set.seed(42)
fabia_2 = fabia(data.fabia, p = 2)
```

```{r}
fabia_2
summary(fabia_2)
plot(fabia_2)
extractPlot(fabia_2)
```

## 3 Biclusters
```{r}
set.seed(42)
fabia_3 = fabia(data.fabia, p = 3)
```
```{r}
fabia_3
```


```{r}
summary(fabia_3)
```
```{r}
plot(fabia_3)
```
```{r}
extractPlot(fabia_3)
```

## 5 Biclusters

```{r}
set.seed(42)
fabia_5 = fabia(data.fabia, p = 5)
```

```{r}
fabia_5
summary(fabia_5)
plot(fabia_5)
extractPlot(fabia_5)
```

## 10 Biclusters

```{r}
set.seed(42)
fabia_10 = fabia(data.fabia, p = 10)
```

```{r}
fabia_10
summary(fabia_10)
plot(fabia_10)
extractPlot(fabia_10)
```

## 15 Biclusters

```{r}
set.seed(42)
fabia_15 = fabia(data.fabia, p = 15)
```

```{r}
fabia_15
summary(fabia_15)
plot(fabia_15)
extractPlot(fabia_15)
```

## 20 Biclusters

```{r}
set.seed(42)
fabia_20 = fabia(data.fabia, p = 20)
```
```{r}
fabia_20
summary(fabia_20)
```

```{r}
plot(fabia_20)
```
```{r}
extractPlot(fabia_20)
```

## 35 Biclusters

```{r}
set.seed(42)
fabia_35 = fabia(data.fabia, p = 35)
```

```{r}
fabia_35
summary(fabia_35)
plot(fabia_35)
extractPlot(fabia_35)
```

## 43 Biclusters (Maximum allowed)

```{r}
set.seed(42)
fabia_43 = fabia(data.fabia, p = 43)
```

```{r}
fabia_43
summary(fabia_43)
plot(fabia_43)
extractPlot(fabia_43)
```
```{r}
fabia_results = list("1bc" = fabia_1,
                     "2bc" = fabia_2,
                     "3bc" = fabia_3,
                     "5bc" = fabia_5,
                     "10bc" = fabia_10,
                     "15bc" = fabia_15,
                     "20bc" = fabia_20,
                     "35bc" = fabia_35,
                     "43bc" = fabia_43)

saveRDS(fabia_results, file = "data/out_models/results_fabia.rds")
```

# 2. Meta Data Exploration

```{r}
data.horizontal = readRDS("data/derived/data_horizontal.rds")
meta.data = data.horizontal$colData.selected
```



## A. Bicluster Loading & Metadata

```{r}
bind.Z.meta = function(fabia.result) {
  Z.df = as.data.frame(t(as.matrix(fabia.result@Z)))
  if (identical(rownames(Z.df), rownames(meta.data))) {
    return(cbind(Z.df, meta.data))
  } else {
    stop("Z and Meta Data Sample names mismatched!")
  }
}
```

### 1 Bicluster

```{r}
#Z.df = as.data.frame(t(as.matrix(fabia_1@Z)))
#identical(rownames(Z.df), rownames(meta.data))
#df.Zmeta = cbind(Z.df, meta.data)
#plot(as.data.frame(t(as.matrix(fabia_1@Z))))
```

Death/Survival time:

```{r}
ggplot(data = bind.Z.meta(fabia_1), aes(x = patientID, y = bicluster1, color = event_time)) +
  geom_point() +
  scale_color_gradient(low = 'green', high = 'red') +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ggtitle("Biscluster scores of patientIDs \n(1-Bicuster FABIA)") +
  theme(plot.title = element_text(hjust = 0.5))
```

Vital Status:

```{r}
ggplot(data = bind.Z.meta(fabia_1), aes(x = patientID, y = bicluster1, color = vital_status)) +
  geom_point() +
  scale_color_manual(values = c("blue","red")) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ggtitle("Biscluster scores of patientIDs \n(1-Bicuster FABIA)") +
  theme(plot.title = element_text(hjust = 0.5))
```


### 2 Biclusters

Death/Survival time:

```{r}
ggplot(data = bind.Z.meta(fabia_2), aes(x = patientID, y = bicluster1, color = event_time)) +
  geom_point() +
  scale_color_gradient(low = 'green', high = 'red') +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ggtitle("Biscluster scores of patientIDs \n(2-Bicusters FABIA)") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data = bind.Z.meta(fabia_2), aes(x = patientID, y = bicluster2, color = event_time)) +
  geom_point() +
  scale_color_gradient(low = 'green', high = 'red') +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ggtitle("Biscluster scores of patientIDs \n(2-Bicusters FABIA)") +
  theme(plot.title = element_text(hjust = 0.5))
```

Vital Status:

```{r}
ggplot(data = bind.Z.meta(fabia_2), aes(x = patientID, y = bicluster1, color = vital_status)) +
  geom_point() +
  scale_color_manual(values = c("blue","red")) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ggtitle("Biscluster scores of patientIDs \n(2-Bicusters FABIA)") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data = bind.Z.meta(fabia_2), aes(x = patientID, y = bicluster2, color = vital_status)) +
  geom_point() +
  scale_color_manual(values = c("blue","red")) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ggtitle("Biscluster scores of patientIDs \n(2-Bicusters FABIA)") +
  theme(plot.title = element_text(hjust = 0.5))
```


### 3 Biclusters

Death/Survival time:

```{r}
ggplot(data = bind.Z.meta(fabia_3), aes(x = patientID, y = bicluster1, color = event_time)) +
  geom_point() +
  scale_color_gradient(low = 'green', high = 'red') +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ggtitle("Biscluster scores of patientIDs \n(3-Bicusters FABIA)") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data = bind.Z.meta(fabia_3), aes(x = patientID, y = bicluster2, color = event_time)) +
  geom_point() +
  scale_color_gradient(low = 'green', high = 'red') +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ggtitle("Biscluster scores of patientIDs \n(3-Bicusters FABIA)") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data = bind.Z.meta(fabia_3), aes(x = patientID, y = bicluster3, color = event_time)) +
  geom_point() +
  scale_color_gradient(low = 'green', high = 'red') +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ggtitle("Biscluster scores of patientIDs \n(3-Bicusters FABIA)") +
  theme(plot.title = element_text(hjust = 0.5))
```


Vital Status:

```{r}
ggplot(data = bind.Z.meta(fabia_3), aes(x = patientID, y = bicluster1, color = vital_status)) +
  geom_point() +
  scale_color_manual(values = c("blue","red")) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ggtitle("Biscluster scores of patientIDs \n(3-Bicusters FABIA)") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data = bind.Z.meta(fabia_3), aes(x = patientID, y = bicluster2, color = vital_status)) +
  geom_point() +
  scale_color_manual(values = c("blue","red")) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ggtitle("Biscluster scores of patientIDs \n(3-Bicusters FABIA)") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data = bind.Z.meta(fabia_3), aes(x = patientID, y = bicluster3, color = vital_status)) +
  geom_point() +
  scale_color_manual(values = c("blue","red")) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  ggtitle("Biscluster scores of patientIDs \n(3-Bicusters FABIA)") +
  theme(plot.title = element_text(hjust = 0.5))
```

# 3. Results comparison

## A. Average information content per bicluster

```{r}
info.content = data.frame(
  biclusters = c(1,2,3,5,10,15,20,35,43),
  info.sum = c(fabia_1@avini[2], 
             fabia_2@avini[3],
             fabia_3@avini[4],
             fabia_5@avini[6],
             fabia_10@avini[11],
             fabia_15@avini[16],
             fabia_20@avini[21],
             fabia_35@avini[36],
             fabia_43@avini[44])
)
```

```{r}
info.content$info.avg = info.content$info.sum/info.content$biclusters
```

```{r}
plot(info.content$biclusters, info.content$info.avg)
#ggplot(data = BCs, aes(x = biclusters))
```

## B. Top n% features of each Bicluster

```{r}
top_x_percent = function(fabia.result, perct = 5, print = T, show.loading = F) {
  quantile = 1 - perct * 0.01
  fabia_L = fabia.result@L
  
  out.list = vector("list", length = ncol(fabia_L))
  names(out.list) = colnames(fabia_L)
  
  for (i in 1:ncol(fabia_L)) {
    L_curr = fabia_L[,i]
    quant = quantile(abs(L_curr), quantile)
    sel = abs(L_curr) >= quant
    out.list[[i]] = L_curr[sel]
    
    if(print) {
      print(colnames(fabia_L)[i])
      if (show.loading) {
        print(L_curr[sel])
      } else {
        print(names(L_curr[sel]))
      }
    }
  }
  
  out.list
}

#boxplot(fabia_1@L)
```

### Top 5%
```{r}
f1_5 = top_x_percent(fabia_1)
f2_5 = top_x_percent(fabia_2)
f3_5 = top_x_percent(fabia_3)
f5_5 = top_x_percent(fabia_5)
f10_5 = top_x_percent(fabia_10)
f15_5 = top_x_percent(fabia_15)
f20_5 = top_x_percent(fabia_20)
f35_5 = top_x_percent(fabia_35)
f43_5 = top_x_percent(fabia_43)
```

```{r}
top_5_perc_result = list("1bc" = f1_5,
                         "2bc" = f2_5,
                         "3bc" = f3_5,
                         "5bc" = f5_5,
                         "10bc" = f10_5,
                         "15bc" = f15_5,
                         "20bc" = f20_5,
                         "35bc" = f35_5,
                         "43bc" = f43_5)

saveRDS(top_5_perc_result, file = "data/sel_features/results_fabia_top_5_percent.rds")
```

### Top 10%

```{r}
f1_10 = top_x_percent(fabia_1, perct = 10)
f2_10 = top_x_percent(fabia_2, perct = 10)
f3_10 = top_x_percent(fabia_3, perct = 10)
f5_10 = top_x_percent(fabia_5, perct = 10)
f10_10 = top_x_percent(fabia_10, perct = 10)
f15_10 = top_x_percent(fabia_15, perct = 10)
f20_10 = top_x_percent(fabia_20, perct = 10)
f35_10 = top_x_percent(fabia_35, perct = 10)
f43_10 = top_x_percent(fabia_43, perct = 10)
```

```{r}
top_10_perc_result = list("1bc" = f1_10,
                         "2bc" = f2_10,
                         "3bc" = f3_10,
                         "5bc" = f5_10,
                         "10bc" = f10_10,
                         "15bc" = f15_10,
                         "20bc" = f20_10,
                         "35bc" = f35_10,
                         "43bc" = f43_10)

saveRDS(top_10_perc_result, file = "data/sel_features/results_fabia_top_10_percent.rds")
```

### Top 25%

```{r}
f1_25 = top_x_percent(fabia_1, perct = 25, print = F)
f2_25 = top_x_percent(fabia_2, perct = 25, print = F)
f3_25 = top_x_percent(fabia_3, perct = 25, print = F)
f5_25 = top_x_percent(fabia_5, perct = 25, print = F)
f10_25 = top_x_percent(fabia_10, perct = 25, print = F)
f15_25 = top_x_percent(fabia_15, perct = 25, print = F)
f20_25 = top_x_percent(fabia_20, perct = 25, print = F)
f35_25 = top_x_percent(fabia_35, perct = 25, print = F)
f43_25 = top_x_percent(fabia_43, perct = 25, print = F)
```

```{r}
top_25_perc_result = list("1bc" = f1_25,
                         "2bc" = f2_25,
                         "3bc" = f3_25,
                         "5bc" = f5_25,
                         "10bc" = f10_25,
                         "15bc" = f15_25,
                         "20bc" = f20_25,
                         "35bc" = f35_25,
                         "43bc" = f43_25)

saveRDS(top_25_perc_result, file = "data/sel_features/results_fabia_top_25_percent.rds")
```

## C. Visualization of top x%

Check max and min value of loadings in order to create symmetrical plots:

```{r}
max_fL = max(max(fabia_1@L),
             max(fabia_2@L),
             max(fabia_3@L),
             max(fabia_5@L),
             max(fabia_10@L),
             max(fabia_15@L),
             max(fabia_20@L),
             max(fabia_35@L),
             max(fabia_43@L)
)
min_fL = min(min(fabia_1@L),
             min(fabia_2@L),
             min(fabia_3@L),
             min(fabia_5@L),
             min(fabia_10@L),
             min(fabia_15@L),
             min(fabia_20@L),
             min(fabia_35@L),
             min(fabia_43@L)
)
c(max_fL, min_fL)
```

The limits of y axis will be (-2.1, 2.1) in order to make them symmetrical.

```{r}
plot.fabia.topx = function(fabia_result, fabia_top5, fabia_top10, fabia_top25, bc, abs.ymax = 2.1) {
  total_bc = fabia_result@p1
  if (bc > total_bc) {
    stop("Requested bicluster number exceeds the total available biclusters.")
  }
  FABIA.L = as.data.frame(fabia_result@L)
  FABIA.L$var = rownames(FABIA.L)
  biclusterx = paste("bicluster", bc, sep = "")
  FABIA.L$top5 = FABIA.L$var %in% names(fabia_top5[[biclusterx]])
  FABIA.L$top10 = FABIA.L$var %in% names(fabia_top10[[biclusterx]])
  FABIA.L$top25 = FABIA.L$var %in% names(fabia_top25[[biclusterx]])
  FABIA.L$top = NA
  for(i in 1:nrow(FABIA.L)) {
    if (FABIA.L[i,]$top25 == F) {
      FABIA.L[i,]$top = NA
    } else if (FABIA.L[i,]$top10 == F) {
      FABIA.L[i,]$top = "Top 11% ~ 25%"
    } else if (FABIA.L[i,]$top5 == F){
      FABIA.L[i,]$top = "Top 06% ~ 10%"
    } else {
      FABIA.L[i,]$top = "Top 05%"
    }
  }
  FABIA.L$dataset = sapply(FABIA.L$var, retrieve.dataset.type)
  #FABIA.L
  ggplot(data = FABIA.L, aes_string(x= "var", y = biclusterx, color = "top", shape="dataset")) +
  geom_point() +
  #geom_text(nudge_y = .04, alpha = 1, size = 3) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  ggtitle(paste(total_bc , "Biclusters")) +
  xlab(biclusterx) +
  ylim(-abs.ymax, abs.ymax)
  
  #if (perc5line) {
  #  perc5 = min(FABIA.L[FABIA.L$top5 == T, biclusterx])
  #  p = p + geom_hline(yintercept = perc5)
  #}
  #p
}

retrieve.dataset.type = function(var.name) {
  if (startsWith(var.name, "hsa")) {
    out = "miRNASeqGene"
  } else if (endsWith(var.name, "RPPA")) {
    out = "RPPAArray"
  } else if (endsWith(var.name, "R2Gn")) {
    out = "RNASeq2GeneNorm"
  } else {
    out = NA
  }
  out
}
```

### 1 Bicluster
```{r}
plot.fabia.topx(fabia_1, f1_5, f1_10, f1_25, 1, abs.ymax = 1)
```

### 2 Biclusters
```{r}
plot.fabia.topx(fabia_2, f2_5, f2_10, f2_25, 1, abs.ymax = 1)
plot.fabia.topx(fabia_2, f2_5, f2_10, f2_25, 2, abs.ymax = 1)
```

### 3 Biclusters
```{r}
plot.fabia.topx(fabia_3, f3_5, f3_10, f3_25, 1, abs.ymax = 1.5)
plot.fabia.topx(fabia_3, f3_5, f3_10, f3_25, 2, abs.ymax = 1.5)
plot.fabia.topx(fabia_3, f3_5, f3_10, f3_25, 3, abs.ymax = 1.5)
```

### 5 Biclusters
```{r}
plot.fabia.topx(fabia_5, f5_5, f5_10, f5_25, 1, abs.ymax = 2)
plot.fabia.topx(fabia_5, f5_5, f5_10, f5_25, 2, abs.ymax = 2)
plot.fabia.topx(fabia_5, f5_5, f5_10, f5_25, 3, abs.ymax = 2)
plot.fabia.topx(fabia_5, f5_5, f5_10, f5_25, 4, abs.ymax = 2)
plot.fabia.topx(fabia_5, f5_5, f5_10, f5_25, 5, abs.ymax = 2)
```

### 10 Biclusters
```{r}
plot.fabia.topx(fabia_10, f10_5, f10_10, f10_25, 1)
plot.fabia.topx(fabia_10, f10_5, f10_10, f10_25, 2)
plot.fabia.topx(fabia_10, f10_5, f10_10, f10_25, 3)
plot.fabia.topx(fabia_10, f10_5, f10_10, f10_25, 4)
plot.fabia.topx(fabia_10, f10_5, f10_10, f10_25, 5)
plot.fabia.topx(fabia_10, f10_5, f10_10, f10_25, 6)
plot.fabia.topx(fabia_10, f10_5, f10_10, f10_25, 7)
plot.fabia.topx(fabia_10, f10_5, f10_10, f10_25, 8)
plot.fabia.topx(fabia_10, f10_5, f10_10, f10_25, 9)
plot.fabia.topx(fabia_10, f10_5, f10_10, f10_25, 10)
```

## D Hard membership by extractBic()

```{r}
select.by.Bic = function(fabia.result, perct = 5, print = T, show.loading = F) {
  fabia_L = fabia.result@L
  
  out.list = vector("list", length = ncol(fabia_L))
  names(out.list) = colnames(fabia_L)
  
  for (i in 1:ncol(fabia_L)) { # Number of biclusters
    L_curr = fabia_L[,i]
    # Biclusters name start with bic entry 2l+1 to 3L
    sel = which(names(L_curr)  %in% extractBic(fabia.result)$bic[2 * ncol(fabia_L) + i][[1]])
    out.list[[i]] = L_curr[sel]
    
    if(print) {
      print(colnames(fabia_L)[i])
      if (show.loading) {
        print(L_curr[sel])
      } else {
        print(names(L_curr[sel]))
      }
    }
  }
  
  out.list
}

#boxplot(fabia_1@L)
```

```{r}
f1_bic = select.by.Bic(fabia_1)
f2_bic = select.by.Bic(fabia_2)
f3_bic = select.by.Bic(fabia_3)
f5_bic = select.by.Bic(fabia_5)
f10_bic = select.by.Bic(fabia_10)
f15_bic = select.by.Bic(fabia_15)
f20_bic = select.by.Bic(fabia_20)
f35_bic = select.by.Bic(fabia_35)
f43_bic = select.by.Bic(fabia_43)
```

```{r}
extracted_by_bic_result = list("1bc" = f1_bic,
                          "2bc" = f2_bic,
                         "3bc" = f3_bic,
                         "5bc" = f5_bic,
                         "10bc" = f10_bic,
                         "15bc" = f15_bic,
                         "20bc" = f20_bic,
                         "35bc" = f35_bic,
                         "43bc" = f43_bic)

saveRDS(extracted_by_bic_result, file = "data/sel_features/results_fabia_extracted_by_bic.rds")
```

plotting:

```{r}
plot.fabia.selbyBic = function(fabia_result, fabia_bic, bc, abs.ymax = 2.1) {
  total_bc = fabia_result@p1
  if (bc > total_bc) {
    stop("Requested bicluster number exceeds the total available biclusters.")
  }
  FABIA.L = as.data.frame(fabia_result@L)
  FABIA.L$var = rownames(FABIA.L)
  biclusterx = paste("bicluster", bc, sep = "")
  FABIA.L$in.bic = FABIA.L$var %in% names(fabia_bic[[biclusterx]])

  FABIA.L$dataset = sapply(FABIA.L$var, retrieve.dataset.type)
  #FABIA.L
  ggplot(data = FABIA.L, aes_string(x= "var", y = biclusterx, color = "in.bic", shape="dataset")) +
  geom_point() +
  #geom_text(nudge_y = .04, alpha = 1, size = 3) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  ggtitle(paste(total_bc , "Biclusters")) +
  xlab(biclusterx) +
  ylim(-abs.ymax, abs.ymax)
  
  #if (perc5line) {
  #  perc5 = min(FABIA.L[FABIA.L$top5 == T, biclusterx])
  #  p = p + geom_hline(yintercept = perc5)
  #}
  #p
}

```

### 1 Bicluster
```{r}
plot.fabia.selbyBic(fabia_1, f1_bic, 1, abs.ymax = 1)
```

### 2 Biclusters
```{r}
plot.fabia.selbyBic(fabia_2, f2_bic, 1, abs.ymax = 1)
plot.fabia.selbyBic(fabia_2, f2_bic, 2, abs.ymax = 1)
```

### 3 Biclusters
```{r}
plot.fabia.selbyBic(fabia_3, f3_bic, 1, abs.ymax = 1.5)
plot.fabia.selbyBic(fabia_3, f3_bic, 2, abs.ymax = 1.5)
plot.fabia.selbyBic(fabia_3, f3_bic, 3, abs.ymax = 1.5)
```

### 5 Biclusters
```{r}
plot.fabia.selbyBic(fabia_5, f5_bic, 1, abs.ymax = 2)
plot.fabia.selbyBic(fabia_5, f5_bic, 2, abs.ymax = 2)
plot.fabia.selbyBic(fabia_5, f5_bic, 3, abs.ymax = 2)
plot.fabia.selbyBic(fabia_5, f5_bic, 4, abs.ymax = 2)
plot.fabia.selbyBic(fabia_5, f5_bic, 5, abs.ymax = 2)
```

### 10 Biclusters
```{r}
plot.fabia.selbyBic(fabia_10, f10_bic, 1)
plot.fabia.selbyBic(fabia_10, f10_bic, 2)
plot.fabia.selbyBic(fabia_10, f10_bic, 3)
plot.fabia.selbyBic(fabia_10, f10_bic, 4)
plot.fabia.selbyBic(fabia_10, f10_bic, 5)
plot.fabia.selbyBic(fabia_10, f10_bic, 6)
plot.fabia.selbyBic(fabia_10, f10_bic, 7)
plot.fabia.selbyBic(fabia_10, f10_bic, 8)
plot.fabia.selbyBic(fabia_10, f10_bic, 9)
plot.fabia.selbyBic(fabia_10, f10_bic, 10)
```
### Manual calculation and display of thres L. 

FABIA does not return the threshold L. We will use the code also used in the ensemble document.

```{r}
get.FABIA.tL = function(FABIA.result, tZ.FABIA = 0.5) {
  
  threasZ.fabia= tZ.FABIA

  tt = 0
  for (i in 1:ncol(L(FABIA.result))) {
    tt = tt + sum(L(FABIA.result)[,i]^2) * sum(Z(FABIA.result)[i,]^2)
  }
  tt = tt/(as.double(nrow(L(FABIA.result))) * as.double(ncol(L(FABIA.result))) * as.double(ncol(Z(FABIA.result))))
  threasL.fabia = sqrt(tt) / threasZ.fabia
  threasL.fabia
}
```

```{r}
print(paste(1, "Bicluster, thresL =", get.FABIA.tL(fabia_1)))
print(paste(2, "Biclusters, thresL =", get.FABIA.tL(fabia_2)))
print(paste(3, "Biclusters, thresL =", get.FABIA.tL(fabia_3)))
print(paste(5, "Biclusters, thresL =", get.FABIA.tL(fabia_5)))
print(paste(10, "Biclusters, thresL =", get.FABIA.tL(fabia_10)))
print(paste(15, "Biclusters, thresL =", get.FABIA.tL(fabia_15)))
print(paste(20, "Biclusters, thresL =", get.FABIA.tL(fabia_20)))
print(paste(35, "Biclusters, thresL =", get.FABIA.tL(fabia_35)))
print(paste(43, "Biclusters, thresL =", get.FABIA.tL(fabia_43)))

```

## E Top n% samples for each Bicluster


```{r}
top_x_percent.s = function(fabia.result, perct = 5, print = F, show.factor = F) {
  quantile = 1 - perct * 0.01
  fabia_Z = t(fabia.result@Z)
  
  out.list = vector("list", length = ncol(fabia_Z))
  names(out.list) = colnames(fabia_Z)
  
  for (i in 1:ncol(fabia_Z)) {
    Z_curr = fabia_Z[,i]
    quant = quantile(abs(Z_curr), quantile)
    sel = abs(Z_curr) >= quant
    out.list[[i]] = Z_curr[sel]
    
    if(print) {
      print(colnames(fabia_Z)[i])
      if (show.factor) {
        print(Z_curr[sel])
      } else {
        print(names(Z_curr[sel]))
      }
    }
  }
  
  out.list
}

#boxplot(fabia_1@L)
```

### Top 5%
```{r}
f1_5.s = top_x_percent.s(fabia_1)
f2_5.s = top_x_percent.s(fabia_2)
f3_5.s = top_x_percent.s(fabia_3)
f5_5.s = top_x_percent.s(fabia_5)
f10_5.s = top_x_percent.s(fabia_10)
f15_5.s = top_x_percent.s(fabia_15)
f20_5.s = top_x_percent.s(fabia_20)
f35_5.s = top_x_percent.s(fabia_35)
f43_5.s = top_x_percent.s(fabia_43)
```

### Top 10%
```{r}
f1_10.s = top_x_percent.s(fabia_1, perct = 10)
f2_10.s = top_x_percent.s(fabia_2, perct = 10)
f3_10.s = top_x_percent.s(fabia_3, perct = 10)
f5_10.s = top_x_percent.s(fabia_5, perct = 10)
f10_10.s = top_x_percent.s(fabia_10, perct = 10)
f15_10.s = top_x_percent.s(fabia_15, perct = 10)
f20_10.s = top_x_percent.s(fabia_20, perct = 10)
f35_10.s = top_x_percent.s(fabia_35, perct = 10)
f43_10.s = top_x_percent.s(fabia_43, perct = 10)
```

### Top 25%
```{r}
f1_25.s = top_x_percent.s(fabia_1, perct = 25)
f2_25.s = top_x_percent.s(fabia_2, perct = 25)
f3_25.s = top_x_percent.s(fabia_3, perct = 25)
f5_25.s = top_x_percent.s(fabia_5, perct = 25)
f10_25.s = top_x_percent.s(fabia_10, perct = 25)
f15_25.s = top_x_percent.s(fabia_15, perct = 25)
f20_25.s = top_x_percent.s(fabia_20, perct = 25)
f35_25.s = top_x_percent.s(fabia_35, perct = 25)
f43_25.s = top_x_percent.s(fabia_43, perct = 25)
```

## C. Visualization of top x% (Max 10 Biclusters)

Check max and min value of loadings in order to create symmetrical plots:

```{r}
max_fZ_all = max(max(fabia_1@Z),
             max(fabia_2@Z),
             max(fabia_3@Z),
             max(fabia_5@Z),
             max(fabia_10@Z),
             max(fabia_15@Z),
             max(fabia_20@Z),
             max(fabia_35@Z),
             max(fabia_43@Z)
)
min_fZ_all = min(min(fabia_1@Z),
             min(fabia_2@Z),
             min(fabia_3@Z),
             min(fabia_5@Z),
             min(fabia_10@Z),
             min(fabia_15@Z),
             min(fabia_20@Z),
             min(fabia_35@Z),
             min(fabia_43@Z)
)
c(max_fZ_all, min_fZ_all)
```


The limits of y axis will be (-5, 5) in order to make them symmetrical.

```{r}
plot.fabia.topx.s = function(fabia_result, fabia_top5, fabia_top10, fabia_top25, bc, abs.ymax = 5) {
  total_bc = fabia_result@p1
  if (bc > total_bc) {
    stop("Requested bicluster number exceeds the total available biclusters.")
  }
  FABIA.Z = as.data.frame(t(fabia_result@Z))
  FABIA.Z$MID = rownames(FABIA.Z)
  biclusterx = paste("bicluster", bc, sep = "")
  FABIA.Z$top5 = FABIA.Z$MID %in% names(fabia_top5[[biclusterx]])
  FABIA.Z$top10 = FABIA.Z$MID %in% names(fabia_top10[[biclusterx]])
  FABIA.Z$top25 = FABIA.Z$MID %in% names(fabia_top25[[biclusterx]])
  FABIA.Z$top = NA
  for(i in 1:nrow(FABIA.Z)) {
    if (FABIA.Z[i,]$top25 == F) {
      FABIA.Z[i,]$top = NA
    } else if (FABIA.Z[i,]$top10 == F) {
      FABIA.Z[i,]$top = "Top 11% ~ 25%"
    } else if (FABIA.Z[i,]$top5 == F){
      FABIA.Z[i,]$top = "Top 06% ~ 10%"
    } else {
      FABIA.Z[i,]$top = "Top 05%"
    }
  }
  #print(FABIA.Z)
  ggplot(data = FABIA.Z, aes_string(x= "MID", y = biclusterx, color = "top", label = "MID")) +
  geom_point() +
  geom_text_repel(data=subset(FABIA.Z, !is.na(top)), alpha = 1, size = 3, fontface = "bold",
                  force = 200, force_pull = 1,
                  max.overlaps = 50) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) +
  ggtitle(paste(total_bc , "Biclusters")) +
  xlab(biclusterx) +
  ylim(-abs.ymax, abs.ymax)
  
  #if (perc5line) {
  #  perc5 = min(FABIA.L[FABIA.L$top5 == T, biclusterx])
  #  p = p + geom_hline(yintercept = perc5)
  #}
  #p
}


```

### 1 Bicluster

```{r}
plot.fabia.topx.s(fabia_1, f1_5.s, f1_10.s, f1_25.s, bc = 1)
```

### 2 Biclusters

```{r}
plot.fabia.topx.s(fabia_2, f2_5.s, f2_10.s, f2_25.s, bc = 1)
plot.fabia.topx.s(fabia_2, f2_5.s, f2_10.s, f2_25.s, bc = 2)
```
### 3 Biclusters

```{r}
plot.fabia.topx.s(fabia_3, f3_5.s, f3_10.s, f3_25.s, bc = 1)
plot.fabia.topx.s(fabia_3, f3_5.s, f3_10.s, f3_25.s, bc = 2)
plot.fabia.topx.s(fabia_3, f3_5.s, f3_10.s, f3_25.s, bc = 3)
```

### 5 Biclusters

```{r}
plot.fabia.topx.s(fabia_5, f5_5.s, f5_10.s, f5_25.s, bc = 1)
plot.fabia.topx.s(fabia_5, f5_5.s, f5_10.s, f5_25.s, bc = 2)
plot.fabia.topx.s(fabia_5, f5_5.s, f5_10.s, f5_25.s, bc = 3)
plot.fabia.topx.s(fabia_5, f5_5.s, f5_10.s, f5_25.s, bc = 4)
plot.fabia.topx.s(fabia_5, f5_5.s, f5_10.s, f5_25.s, bc = 5)
```

### 10 Biclusters

```{r}
plot.fabia.topx.s(fabia_10, f10_5.s, f10_10.s, f10_25.s, bc = 1)
plot.fabia.topx.s(fabia_10, f10_5.s, f10_10.s, f10_25.s, bc = 2)
plot.fabia.topx.s(fabia_10, f10_5.s, f10_10.s, f10_25.s, bc = 3)
plot.fabia.topx.s(fabia_10, f10_5.s, f10_10.s, f10_25.s, bc = 4)
plot.fabia.topx.s(fabia_10, f10_5.s, f10_10.s, f10_25.s, bc = 5)
plot.fabia.topx.s(fabia_10, f10_5.s, f10_10.s, f10_25.s, bc = 6)
plot.fabia.topx.s(fabia_10, f10_5.s, f10_10.s, f10_25.s, bc = 7)
plot.fabia.topx.s(fabia_10, f10_5.s, f10_10.s, f10_25.s, bc = 8)
plot.fabia.topx.s(fabia_10, f10_5.s, f10_10.s, f10_25.s, bc = 9)
plot.fabia.topx.s(fabia_10, f10_5.s, f10_10.s, f10_25.s, bc = 10)
```
