---
title: "MFA"
author: "John Zhuang"
date: "2024-05-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(FactoMineR)
library(factoextra)
library(ggplot2)
library(scales)
```

```{r}
data.horizontal = readRDS("data/derived/data_horizontal_joined.rds")
data.MFA = data.horizontal
```


# 1. MFA



## MFA with only numerical datasets

```{r}
set.seed(42)
MFA.numeric.only = MFA(data.MFA[,1:701],
                 group = c(470, 198, 33),
                 type = c("c", "c", "c"),
                 ncp = 10,
                 name.group = c("miRNASeqGene", "RNASeq2GeneNorm", "RPPAArray"))
```

## MFA Including some categorical metadata (Used only as a reference)

```{r}
set.seed(42)
MFA.numeric.withd = MFA(data.MFA[,c(1:701, 998, 1001:1005)], 
                 group = c(470, 198, 33, 6),
                 type = c("c", "c", "c", "n"),
                 ncp = 10,
                 name.group = c("miRNASeqGene", "RNASeq2GeneNorm", "RPPAArray", "Demographics"))
names(data.MFA[,c(998, 1001:1005)])
```

```{r}
MFA.models = list("n.only" = MFA.numeric.only,
                  "w.meta" = MFA.numeric.withd)
saveRDS(MFA.models, file = "data/out_models/results_mfa.rds")
```


# 2. Results Representation

## A. Samples and Variables contributions

```{r}
MFA.numeric.only$eig
```

```{r}
MFA.results.samples = get_mfa_ind(MFA.numeric.only)
MFA.results.samples$contrib
```

```{r}
MFA.results.variables = get_mfa_var(MFA.numeric.only)
MFA.results.variables$contrib
MFA.results.variables$cos2
```
```{r}
MFA.results.axes = get_mfa_partial_axes(MFA.numeric.only)
MFA.results.axes$contrib
```

## B. Eigenvalues

```{r}
eigen.MFA = get_eigenvalue(MFA.numeric.only)
head(eigen.MFA, 10)
fviz_screeplot(MFA.numeric.only,
               ncp = 10,
               addlabels = T,
               main = "Top 10 Dimensions by Variance Explained")
```

## C. Variable groups

```{r}
group.MFA = get_mfa_var(MFA.numeric.only, "group")
fviz_mfa_var(MFA.numeric.only, "group")

fviz_contrib(MFA.numeric.only, "group", axes = 1)
fviz_contrib(MFA.numeric.only, "group", axes = 2)
fviz_contrib(MFA.numeric.only, "group", axes = 3)
fviz_contrib(MFA.numeric.only, "group", axes = 4)
fviz_contrib(MFA.numeric.only, "group", axes = 5)
fviz_contrib(MFA.numeric.only, "group", axes = 6)
fviz_contrib(MFA.numeric.only, "group", axes = 7)
fviz_contrib(MFA.numeric.only, "group", axes = 8)
fviz_contrib(MFA.numeric.only, "group", axes = 9)
fviz_contrib(MFA.numeric.only, "group", axes = 10)
```

### Print contribution percentage of top 3 dimensions:

```{r}
fviz_contrib(MFA.numeric.only, "group", axes = 1)$data
fviz_contrib(MFA.numeric.only, "group", axes = 2)$data
fviz_contrib(MFA.numeric.only, "group", axes = 3)$data
```


## D. Contribution of Independent Quantitative Variables

```{r}
quant.var.MFA = get_mfa_var(MFA.numeric.only, "quanti.var")

fviz_mfa_var(MFA.numeric.only, "quanti.var",
             col.var.sup = "violet", repel = TRUE,
             geom = "point", legend = "bottom")

fviz_contrib(MFA.numeric.only, choice = "quanti.var", axes = 1, top = 30)
fviz_contrib(MFA.numeric.only, choice = "quanti.var", axes = 2, top = 30, palette= hue_pal()(3)[2:3])
fviz_contrib(MFA.numeric.only, choice = "quanti.var", axes = 3, top = 30, palette= hue_pal()(3)[2:3])
fviz_contrib(MFA.numeric.only, choice = "quanti.var", axes = 4, top = 30)
fviz_contrib(MFA.numeric.only, choice = "quanti.var", axes = 5, top = 30)
fviz_contrib(MFA.numeric.only, choice = "quanti.var", axes = 6, top = 30)
fviz_contrib(MFA.numeric.only, choice = "quanti.var", axes = 7, top = 30)
fviz_contrib(MFA.numeric.only, choice = "quanti.var", axes = 8, top = 30)
fviz_contrib(MFA.numeric.only, choice = "quanti.var", axes = 9, top = 30)
fviz_contrib(MFA.numeric.only, choice = "quanti.var", axes = 10, top = 30)
```

## E. Plot of individuals

```{r}
MFA.ind = get_mfa_ind(MFA.numeric.only)

#fviz_mfa_ind(MFA.numeric.only, col.ind = "cos2", 
#             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
#             repel = TRUE)

fviz_mfa_ind(MFA.numeric.only,
             geom = "point")

fviz_mfa_var(MFA.numeric.only,
             geom = "point")

fviz_mfa_ind(MFA.numeric.withd, 
             habillage = "vital_status",
             palette = c("blue", "red"),
             addEllipses = TRUE, ellipse.type = "confidence", 
             repel = TRUE, # Avoid text overlapping
             geom = "point"
             ) 

fviz_ellipses(MFA.numeric.withd, 
             c("vital_status", "gender", "radiation_therapy",
               "pathologic_stage", "pathology_T_stage", "pathology_N_stage"),
             repel = TRUE, # Avoid text overlapping
             palette = "simpsons",
             geom = "point"
             ) 

#fviz_mfa_ind(MFA.numeric.withd, 
#             habillage = "event_time",
#             geom = "point"
#             ) 
```

###  Manual Plotting of 1st and 2nd dimensions

```{r}
dim123 = as.data.frame(MFA.results.samples$coord[,1:3])
vital_status = data.MFA$vital_status
all(rownames(dim123) == rownames(vital_status))
merged.plotdata = cbind(dim123, vital_status)

ggplot(data = merged.plotdata, aes(x=Dim.1, y = Dim.2, color = vital_status)) +
  geom_point() +
  stat_ellipse(level = 0.1) +
  stat_ellipse(level = 0.001, geom = "point") +
  scale_color_manual(values = c("blue", "red")) +
  ggtitle("Vital Status of samples, MFA Dimensions 1 & 2")
```
```{r}
ggplot(data = merged.plotdata, aes(x=Dim.2, y = Dim.3, color = vital_status)) +
  geom_point() +
  stat_ellipse(level = 0.1) +
  stat_ellipse(level = 0.001, geom = "point") +
  scale_color_manual(values = c("blue", "red")) +
  ggtitle("Vital Status of samples, MFA Dimensions 2 & 3")
```


## F. Partial Individuals

```{r}
fviz_mfa_ind(MFA.numeric.only, partial = "all") 
fviz_mfa_ind(MFA.numeric.only, partial = c("TCGA-OR-A5LT", "TCGA-PK-A5H9")) 
```

## G. Partial Axes

```{r}
fviz_mfa_axes(MFA.numeric.only)
```

# 3. Save MFA variable contribution to RDS file

We only keep dimensions with eigenvalues >= 1. 

In this case, it's the first 3rd dimensions.

```{r}
eig.values = get_eigenvalue(MFA.numeric.only)
eig.values[which(eig.values[,1] > 1),]
```

Extract and order first 3 dimensions by variable contribution
```{r}
MFA.variables.contrib = MFA.results.variables$contrib
contrib.dim1 = MFA.variables.contrib[order(MFA.variables.contrib[,1], decreasing = T), 1]
contrib.dim2 = MFA.variables.contrib[order(MFA.variables.contrib[,2], decreasing = T), 2]
contrib.dim3 = MFA.variables.contrib[order(MFA.variables.contrib[,3], decreasing = T), 3]
```

Only keep variables with a greater than mean contribution.
```{r}
mean.contrib = 100/dim(MFA.variables.contrib)[1]
contrib.dim1.sel = contrib.dim1[which(contrib.dim1 > mean.contrib)]
contrib.dim2.sel = contrib.dim2[which(contrib.dim2 > mean.contrib)]
contrib.dim3.sel = contrib.dim3[which(contrib.dim3 > mean.contrib)]
```



```{r}
results.mfa = list("Dim1" = contrib.dim1.sel,
                   "Dim2" = contrib.dim2.sel,
                   "Dim3" = contrib.dim3.sel)
saveRDS(results.mfa, file = "data/sel_features/results_mfa_selected.rds")
```

