---
title: "MOFA seed difference checker"
author: "John Zhuang"
date: "2024-06-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MOFA2)
library(utils)
library(stats)
library(ggplot2)
```

```{r}
data.vertical = readRDS("data/derived/data_vertical_numeric_only.rds")
```

# 0. Overview

This document aims to check if MOFA returns different results with different seeds under the same setting.

# 1. MOFA Training

## A. Check again to reassure if colnames are identical before analysis
```{r}
#
identical(colnames(data.vertical$RPPAArray), colnames(data.vertical$RNASeq2GeneNorm))
identical(colnames(data.vertical$RNASeq2GeneNorm), colnames(data.vertical$miRNASeqGene))
```
## B. Create MOFA object

```{r}
matrix.list = list("view_mRNA" = as.matrix(data.vertical$miRNASeqGene),
                   "view_R2Gn" = as.matrix(data.vertical$RNASeq2GeneNorm),
                   "view_RPPA" = as.matrix(data.vertical$RPPAArray))
```

```{r}
data.mofa = create_mofa(matrix.list)
```
```{r}
plot_data_overview(data.mofa)
```

## C. Add metadata to the model

```{r}
data.horizontal = readRDS("data/derived/data_horizontal.rds")
meta.data = data.horizontal$colData.selected
```

```{r}
#Rename patientID to sample for MOFA
names(meta.data)[1] = "sample"

samples_metadata(data.mofa) = meta.data
head(samples_metadata(data.mofa))
```

## D. Model training parameters
```{r}
data.opts = get_default_data_options(data.mofa)
data.opts
model.opts = get_default_model_options(data.mofa)
model.opts
training.opts = get_default_training_options(data.mofa)
training.opts
```
The output warns the default 15 factors to be too many for the sample size.
Thus, we reduce it to the 10 (maximum recommended):

```{r}
model.opts_10 = get_default_model_options(data.mofa)
model.opts_10$num_factors = 10
```

We will also do 1, 2, 3 and 5 factors in order to compare results with those FABIA.

```{r}
model.opts_1 = get_default_model_options(data.mofa)
model.opts_2 = get_default_model_options(data.mofa)
model.opts_3 = get_default_model_options(data.mofa)
model.opts_5 = get_default_model_options(data.mofa)

model.opts_1$num_factors = 1
model.opts_2$num_factors = 2
model.opts_3$num_factors = 3
model.opts_5$num_factors = 5
```

## E. Set 5 different seeds.

```{r}
# Generate one seed per run
set.seed(42)
seeds = sample.int(100, 5)
seeds
```

## F. Train the model (10 factors).

```{r}
data.mofa.ready_10 = prepare_mofa(
  object = data.mofa,
  data_options = data.opts,
  model_options = model.opts_10,
  training_options = training.opts
)
```

### Attempt 1
```{r, cache=F}
outfile = "data/temp/mofat1.hdf5"
set.seed(seeds[1])
MOFA.t1 <- run_mofa(data.mofa.ready_10, outfile, use_basilisk=T)
```

### Attempt 2
```{r, cache=F}
outfile = "data/temp/mofat2.hdf5"
set.seed(seeds[2])
MOFA.t2 <- run_mofa(data.mofa.ready_10, outfile, use_basilisk=T)
```

### Attempt 3
```{r, cache=F}
outfile = "data/temp/mofat3.hdf5"
set.seed(seeds[3])
MOFA.t3 <- run_mofa(data.mofa.ready_10, outfile, use_basilisk=T)
```

### Attempt 4
```{r, cache=F}
outfile = "data/temp/mofat4.hdf5"
set.seed(seeds[4])
MOFA.t4 <- run_mofa(data.mofa.ready_10, outfile, use_basilisk=T)
```

### Attempt 5
```{r, cache=F}
outfile = "data/temp/mofat5.hdf5"
set.seed(seeds[5])
MOFA.t5 <- run_mofa(data.mofa.ready_10, outfile, use_basilisk=T)
```

# 2. Compare Results

```{r}
factors1 = get_factors(MOFA.t1, as.data.frame = T)
weights1 = get_weights(MOFA.t1, as.data.frame = T)

factors2 = get_factors(MOFA.t2, as.data.frame = T)
weights2 = get_weights(MOFA.t2, as.data.frame = T)

factors3 = get_factors(MOFA.t3, as.data.frame = T)
weights3 = get_weights(MOFA.t3, as.data.frame = T)

factors4 = get_factors(MOFA.t4, as.data.frame = T)
weights4 = get_weights(MOFA.t4, as.data.frame = T)

factors5 = get_factors(MOFA.t5, as.data.frame = T)
weights5 = get_weights(MOFA.t5, as.data.frame = T)
```

```{r}
all(weights1 == weights2)
all(weights2 == weights3)
all(weights3 == weights4)
all(weights4 == weights5)

all(factors1 == factors2)
all(factors2 == factors3)
all(factors3 == factors4)
all(factors4 == factors5)
```
It can be seen that all MOFA results are equal regardless of the seed set in R.
