---
title: "FABIA and MOFA results Comparison"
author: "John Zhuang"
date: "2024-06-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(ggpubr)
#library(grid)
#library(gridExtra)
library(ggrepel)
library(dplyr)
library(RColorBrewer)
```

# 0. Preparation

## A. Load FABIA and MOFA results

```{r}
FABIA.top5= readRDS(file = "data/sel_features/results_fabia_top_5_percent.rds")
FABIA.top10= readRDS(file = "data/sel_features/results_fabia_top_10_percent.rds")
FABIA.top25= readRDS(file = "data/sel_features/results_fabia_top_25_percent.rds")
FABIA.Bic = readRDS(file = "data/sel_features/results_fabia_extracted_by_bic.rds")

MOFA.top5= readRDS(file = "data/sel_features/results_mofa_top_5_percent.rds")
MOFA.top10= readRDS(file = "data/sel_features/results_mofa_top_10_percent.rds")
MOFA.top25= readRDS(file = "data/sel_features/results_mofa_top_25_percent.rds")
MOFA.Bic = readRDS(file = "data/sel_features/results_mofa_extracted_by_bic.rds")
```

## B. Join FABIA and MOFA top X% loadings

```{r}
merge.FABIA.MOFA.results = function(FABIA.results, MOFA.results, total_bcf, bc, f = bc, complete = F) {
  if (bc > total_bcf) {
    stop("Requested bicluster number exceeds the total available biclusters.")
  }
  if (!(is.na(f)) & f > total_bcf) {
    stop("Requested factor number exceeds the total available factors.")
  }
  
  xbc = paste(total_bcf, "bc", sep = "")
  biclusterx = paste("bicluster", bc, sep = "")
  xf = paste(total_bcf, "f", sep = "")
  Factorx = paste("Factor", f, sep = "")
  FABIA.loadings = FABIA.results[[xbc]][[biclusterx]]
  MOFA.weights = MOFA.results[[xf]][[Factorx]]
  FABIA.df = as.data.frame(FABIA.loadings)
  MOFA.df = as.data.frame(MOFA.weights)
  FABIA.df$var = rownames(FABIA.df)
  MOFA.df$var = rownames(MOFA.df)
  #FABIA.df$FABIA.absL = abs(FABIA.df$FABIA.loadings) # Absolute values dropped: Calculated now by ggplot
  #MOFA.df$MOFA.absW = abs(MOFA.df$MOFA.weights)
  out = full_join(FABIA.df, MOFA.df, by = "var")
  if (complete) {
    out = out[complete.cases(out), ]
  }
  out
}
#fm.1.1 = full_join(FABIA.top5$`1bc`$bicluster1, MOFA.top5$`1bc`$bicluster1, by = )
```

```{r}
# Old code to check number of shared features across FABIA and MOFA results.
# Deprecated as now FABIA & MOFA results are plotted in grids
check.FABIA.MOFA.overlap = function(FABIA.results, MOFA.results, total_bcf, against.plot = F) {
  out.matrix = matrix(, nrow = total_bcf, ncol = total_bcf)
  for(f in 1:total_bcf) {
    for(bc in 1:total_bcf) {
      curr.fm = merge.FABIA.MOFA.results(FABIA.results, MOFA.results, total_bcf, bc = bc, f = f)
      overlap = sum(complete.cases(curr.fm))
      #If using this function to check against plots, overlaps <= 2 are treated as no overlap
      if (against.plot & overlap < 3) { 
        overlap = 0
      }
      out.matrix[f,bc] = overlap
    } 
  }
  out.matrix
}
```


```{r}
# Extract data omics category and add color for plotting
retrieve.dataset.type = function(var.name) {
  if (startsWith(var.name, "hsa")) {
    out = "miRNASeqGene"
  } else if (endsWith(var.name, "RPPA")) {
    out = "RPPAArray"
  } else if (endsWith(var.name, "R2Gn")) {
    out = "RNASeq2GeneNorm"
  } else {
    out = NA
  }
  out
}

data.colors = setNames(c("red", "forestgreen", "blue"), c("miRNASeqGene", "RPPAArray", "RNASeq2GeneNorm"))
```


```{r}
#Default data for generating a blank grid
default.data = data.frame(var = c("ph1", "ph2"),
                          FABIA.loadings = c(.8, 1.2),
                          MOFA.weights = c(.8, 1.2),
                          dataset = c("placeholder", "placeholder"))
```

```{r}
plot.mf = function(plot.data) {
  #plot.data = merged.mf.results[complete.cases(merged.mf.results), ] # Select complete cases only to avoid warnings, now deprecated as complete cases selection is done in FABIA-MOFA result merge.
  plot.data$dataset = sapply(plot.data$var, retrieve.dataset.type)
  if (nrow(plot.data) >= 3) {   # Minimum of 3 entries required to plot
    p = ggplot(plot.data, aes(x=FABIA.loadings, y = MOFA.weights)) +
    geom_point(show.legend = F, aes(x = abs(FABIA.loadings), y = abs(MOFA.weights),color = dataset)) +
    stat_cor(method = "spearman", label.y.npc="top", label.x.npc = "left") +
    theme(
      axis.title = element_blank()
    ) +
    scale_color_manual(values = data.colors) #+
    #xlim(min(abs(plot.data$FABIA.loadings)), max(abs(plot.data$FABIA.loadings)))
    #geom_smooth(method = "lm") +
    
    #geom_text(nudge_y = .01, alpha = 0.4)
  } else { # Print blank default plot if 2 or fewer shared variables.
    #print(1)
    p = ggplot(data = default.data, aes(x=FABIA.loadings, y = MOFA.weights, color = dataset)) +
    theme(
      axis.title = element_blank()
    ) 
  }
  p
}
```

```{r}
#Plots without Spearman Correlation. Sets the range of x and y values to the min and max absolute values to allow better viewing.
plot.mf.nospear = function(plot.data) {
  #plot.data = merged.mf.results[complete.cases(merged.mf.results), ] # Select complete cases only to avoid warnings, now deprecated as complete cases selection is done in FABIA-MOFA result merge.
  plot.data$dataset = sapply(plot.data$var, retrieve.dataset.type)
  if (nrow(plot.data) >= 3) {   # Minimum of 3 entries required to plot
    p = ggplot(plot.data, aes(x=FABIA.loadings, y = MOFA.weights)) +
    geom_point(show.legend = F, aes(x = abs(FABIA.loadings), y = abs(MOFA.weights),color = dataset)) +
    #stat_cor(method = "spearman", label.y.npc="top", label.x.npc = "left") +
    theme(
      axis.title = element_blank()
    ) +
    scale_color_manual(values = data.colors) +
    xlim(min(abs(plot.data$FABIA.loadings)), max(abs(plot.data$FABIA.loadings))) +
    ylim(min(abs(plot.data$MOFA.weights)), max(abs(plot.data$MOFA.weights)))
    #geom_smooth(method = "lm") +
    
    #geom_text(nudge_y = .01, alpha = 0.4)
  } else { # Print blank default plot if 2 or fewer shared variables.
    #print(1)
    p = ggplot(data = default.data, aes(x=FABIA.loadings, y = MOFA.weights, color = dataset)) +
    theme(
      axis.title = element_blank()
    ) 
  }
  p
}
```

```{r}
plot.mf.grid = function(FABIA.results, MOFA.results, total_bcf,
                        f_end = total_bcf, bc_end = total_bcf, f_start = 1, bc_start = 1, plot.spearman = T) {
  plot.list = vector('list', (f_end - f_start+1) * (bc_end - bc_start + 1))
  #spearman.list = vector('character', (f_end - f_start+1) * (bc_end - bc_start + 1))
  for(bc in bc_start:bc_end) {
    for(f in f_start:f_end) {
      curr.mf = merge.FABIA.MOFA.results(FABIA.results, MOFA.results, total_bcf, bc = bc, f = f, complete = T)
      if(plot.spearman) {
        plot.i = plot.mf(curr.mf)
      } else {
        plot.i = plot.mf.nospear(curr.mf)
      }
      plot.list[[(bc_end - bc_start+1)*(f-f_start)+(bc-bc_start+1)]] = plot.i
      #if (sum(complete.cases(curr.mf)) >= 3) {
      #  spearman.data = curr.mf[complete.cases(curr.mf), ]
      #  spearman.test = cor.test(x = spearman.data$FABIA.loadings,
      #                           y = spearman.data$MOFA.weights,
      #                           method="spearman")
      #  spearman.list[[(bc_end - bc_start+1)*(f-f_start)+(bc-bc_start+1)]] == paste("Rho:", spearman.test$estimate, "P-value:", spearman.test$estimate)
      #}
    }
  }
  #plot.list
  f = ggarrange(plotlist = plot.list,
            common.legend = T,
            legend = "right",
            #labels = spearman.list,
            nrow = f_end-f_start+1, ncol = bc_end-bc_start+1)
  #grid.arrange(
  #  arrangeGrob(grobs = plot.list, ncol = total_bcf, nrow = total_bcf,
  #              bottom=textGrob("FABIA Loadings (Absolute Value)", gp=gpar(fontface="bold", col="red", fontsize=15)), 
  #              left=textGrob("MOFA Weights, (Absolute Value)", gp=gpar(fontface="bold", col="blue", fontsize=15), rot=90)),
  #)
  bcf.text = paste(total_bcf, "Biclusters/Factors")
  top.annotate = text_grob(bcf.text, color = "black", face = "bold", size = 20)
  if (any(bc_end != total_bcf, f_end != total_bcf, bc_start != 1, f_start != 1)) {
    bcf.text = paste(bcf.text, "\n", "(Biclusters", bc_start, "to", bc_end, ",",
                   "Factors", f_start, "to", f_end,")")
    top.annotate = text_grob(bcf.text, color = "black", face = "bold", size = 10)
  }
  annotate_figure(f,
                  top = top.annotate,
                  bottom = text_grob("FABIA Loadings (Absolute Value)", color = "blue", face = "bold", size = 15),
                  left = text_grob("MOFA Weights, (Absolute Value)", color = "forestgreen", face = "bold", size = 15, rot = 90))
  #print(spearman.list)
}
```

```{r}
# Plot only one pair of bicluster/factor with added labels support (for individual plots exploration)
plot.mf.one = function(FABIA.results, MOFA.results, total_bcf, bc, f) {
  plot.data = merge.FABIA.MOFA.results(FABIA.results, MOFA.results, total_bcf, bc, f, complete = T)
  
  plot.data$dataset = sapply(plot.data$var, retrieve.dataset.type)
  if (nrow(plot.data) >= 3) {   # Minimum of 3 entries required to plot
    p = ggplot(plot.data, aes(x=FABIA.loadings, y = MOFA.weights, label = var)) +
    geom_point(show.legend = F, aes(color = dataset)) +
    stat_cor(method = "spearman", label.y.npc="bottom", label.x.npc = 0.8) +
    scale_color_manual(values = data.colors) +
    geom_text_repel(data=plot.data, alpha = 1, size = 3,
                  force = 300, force_pull = 1,
                  max.overlaps = 10) 
  } else { # Print blank default plot if 2 or fewer shared variables.

    p = ggplot(data = default.data, aes(x=FABIA.loadings, y = MOFA.weights, color = dataset)) 
  }
  p = p + theme(
      axis.title = element_blank()
    ) 
  oneplot.text = paste("Bicluster", bc, "&", "Factor", f, "\n(Of", total_bcf, "Biclusters/Factors)")
  top.annotate = text_grob(oneplot.text, color = "black", face = "bold", size = 10)
  annotate_figure(p,
                  top = top.annotate,
                  bottom = text_grob("FABIA Loadings", color = "blue", face = "bold", size = 15),
                  left = text_grob("MOFA Weights", color = "forestgreen", face = "bold", size = 15, rot = 90))
}
```

```{r, eval = F}
# Old code for manually computing Spearman correlation.
# Deprecated in favor of plotting it directly on the plot.
spearman.FABIA.MOFA = function(FABIA.results, MOFA.results, total_bcf) {
  out.matrix = matrix(, nrow = total_bcf, ncol = total_bcf)
  for(f in 1:total_bcf) {
    for(bc in 1:total_bcf) {
      curr.fm = merge.FABIA.MOFA.results(FABIA.results, MOFA.results, total_bcf, bc = bc, f = f)
      spearman.data = curr.fm[complete.cases(curr.fm), ]
      #print(spearman.data)
      #out.matrix[bc,f] = cor.test(x = spearman.data$FABIA.absL, y = spearman.data$MOFA.absW, method="spearman")
      test = cor.test(x = spearman.data$FABIA.loadings, y = spearman.data$MOFA.weights, method="spearman")
    } 
  }
  #out.matrix
  test
}
```


# 1. FABIA-MOFA variables comparison (Extracted by extractBic() L threashold)

```{r}
check.FABIA.MOFA.overlap(FABIA.Bic, MOFA.Bic, 1, against.plot = T)
```

```{r, fig.width=8, fig.height=8}
plot.mf.grid(FABIA.Bic, MOFA.Bic, 1)
```
```{r}
check.FABIA.MOFA.overlap(FABIA.Bic, MOFA.Bic, 2, against.plot = T)
```

```{r, fig.width=8, fig.height=8}
plot.mf.grid(FABIA.Bic, MOFA.Bic, 2)
plot.mf.grid(FABIA.Bic, MOFA.Bic, 2, plot.spearman = F)
```
```{r}
check.FABIA.MOFA.overlap(FABIA.Bic, MOFA.Bic, 3, against.plot = T)
```

```{r, fig.width=8, fig.height=8}
plot.mf.grid(FABIA.Bic, MOFA.Bic, 3)
plot.mf.grid(FABIA.Bic, MOFA.Bic, 3, plot.spearman = F)
```
```{r}
check.FABIA.MOFA.overlap(FABIA.Bic, MOFA.Bic, 5, against.plot = T)
```



```{r, fig.width=10, fig.height=10}
#dev.new(width = 10, height = 10, unit = "in", noRStudioGD = T)
plot.mf.grid(FABIA.Bic, MOFA.Bic, 5)
plot.mf.grid(FABIA.Bic, MOFA.Bic, 5, plot.spearman = F)
```

```{r}
check.FABIA.MOFA.overlap(FABIA.Bic, MOFA.Bic, 5, against.plot = T)
```

```{r, fig.width=20, fig.height=20}
#dev.new(width = 20, height = 20, unit = "in", noRStudioGD = T)
plot.mf.grid(FABIA.Bic, MOFA.Bic, 10)
plot.mf.grid(FABIA.Bic, MOFA.Bic, 10, plot.spearman = F)
```
## 1B. Individual Plots Worth Exploring:

```{r, fig.width=8, fig.height=8}
plot.mf.one(FABIA.Bic, MOFA.Bic, 1, 1, 1)
```
```{r, fig.width=8, fig.height=8}
plot.mf.one(FABIA.Bic, MOFA.Bic, 3, 2, 1)
```
```{r, fig.width=8, fig.height=8}
plot.mf.one(FABIA.Bic, MOFA.Bic, 3, 3, 1)
```

```{r, fig.width=8, fig.height=8}
plot.mf.one(FABIA.Bic, MOFA.Bic, 5, 1, 2)
plot.mf.one(FABIA.Bic, MOFA.Bic, 5, 2, 2)
plot.mf.one(FABIA.Bic, MOFA.Bic, 5, 2, 3)
plot.mf.one(FABIA.Bic, MOFA.Bic, 5, 3, 2)
plot.mf.one(FABIA.Bic, MOFA.Bic, 5, 5, 2)
```


```{r, fig.width=8, fig.height=8}
plot.mf.one(FABIA.Bic, MOFA.Bic, 10, 1, 2)
plot.mf.one(FABIA.Bic, MOFA.Bic, 10, 2, 2)
plot.mf.one(FABIA.Bic, MOFA.Bic, 10, 3, 2)
plot.mf.one(FABIA.Bic, MOFA.Bic, 10, 9, 2)
plot.mf.one(FABIA.Bic, MOFA.Bic, 10, 10, 2)
plot.mf.one(FABIA.Bic, MOFA.Bic, 10, 7, 4)
plot.mf.one(FABIA.Bic, MOFA.Bic, 10, 9, 4)
plot.mf.one(FABIA.Bic, MOFA.Bic, 10, 10, 4)
```

Export overlaping Biclusters with linear combination of bicluster and factor
```{r}
fabia.mofa = list("b1f1.1" = merge.FABIA.MOFA.results(FABIA.Bic, MOFA.Bic, 1, 1, 1, complete = T),
                  "b1f2.10" = merge.FABIA.MOFA.results(FABIA.Bic, MOFA.Bic, 10, 1, 2, complete = T))
saveRDS(fabia.mofa, file = "data/sel_features/fabia_mofa_overlap_of_interest.rds")
```


Deprecated, kept for completeness:

# 2. FABIA-MOFA variables comparison (Top 5%)

```{r, fig.width=8, fig.height=8}
plot.mf.grid(FABIA.top5, MOFA.top5, 1)
```


```{r, fig.width=8, fig.height=8}
plot.mf.grid(FABIA.top5, MOFA.top5, 2)
```


```{r, fig.width=8, fig.height=8}
plot.mf.grid(FABIA.top5, MOFA.top5, 3)
```


```{r, fig.width=10, fig.height=10}
#dev.new(width = 10, height = 10, unit = "in", noRStudioGD = T)
plot.mf.grid(FABIA.top5, MOFA.top5, 5)
```

```{r, eval=F}
plot.mf.grid(FABIA.top5, MOFA.top5, 10, 5, 5, 1, 1)
plot.mf.grid(FABIA.top5, MOFA.top5, 10, 5, 10, 1, 6)
plot.mf.grid(FABIA.top5, MOFA.top5, 10, 10, 5, 6, 1)
plot.mf.grid(FABIA.top5, MOFA.top5, 10, 10, 10, 6, 6)
```

```{r, fig.width=20, fig.height=20}
#dev.new(width = 20, height = 20, unit = "in", noRStudioGD = T)
plot.mf.grid(FABIA.top5, MOFA.top5, 10)
```

## 2B. Individual Plots Worth Exploring:

```{r, fig.width=8, fig.height=8}
plot.mf.one(FABIA.top5, MOFA.top5, 1, 1, 1)
```

```{r, fig.width=8, fig.height=8}
plot.mf.one(FABIA.top5, MOFA.top5, 5, 2, 2)
```
```{r, fig.width=8, fig.height=8}
plot.mf.one(FABIA.top5, MOFA.top5, 10, 1, 2)
plot.mf.one(FABIA.top5, MOFA.top5, 10, 2, 2)

```

# 3. FABIA-MOFA variables comparison (Top 10%)

```{r}
plot.mf.grid(FABIA.top10, MOFA.top10, 1)
```

```{r}
plot.mf.grid(FABIA.top10, MOFA.top10, 2)
```

```{r}
plot.mf.grid(FABIA.top10, MOFA.top10, 3)
```

```{r, fig.width=10, fig.height=10}
#dev.new(width = 10, height = 10, unit = "in", noRStudioGD = T)
plot.mf.grid(FABIA.top10, MOFA.top10, 5)
```


```{r, fig.width=20, fig.height=20}
#dev.new(width = 20, height = 20, unit = "in", noRStudioGD = T)
plot.mf.grid(FABIA.top10, MOFA.top10, 10)
```

# 4. FABIA-MOFA variables comparison (Top 25%)

```{r}
plot.mf.grid(FABIA.top25, MOFA.top25, 1)
```

```{r}
plot.mf.grid(FABIA.top25, MOFA.top25, 2)
```

```{r}
plot.mf.grid(FABIA.top25, MOFA.top25, 3)
```

```{r, fig.width=10, fig.height=10}
#dev.new(width = 10, height = 10, unit = "in", noRStudioGD = T)
plot.mf.grid(FABIA.top25, MOFA.top25, 5)
```

```{r, fig.width=20, fig.height=20}
#dev.new(width = 20, height = 20, unit = "in", noRStudioGD = T)
plot.mf.grid(FABIA.top25, MOFA.top25, 10)
```

