---
title: "Ensemble Methods"
author: "John Zhuang"
date: "2024-05-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(fabia)
library(MOFA2)
#library(biclust)
library(superbiclust)
library(ggplot2)
```

```{r}
data.vertical = readRDS("data/derived/data_vertical_numeric_only.rds")
```

# 0. Preparation

## a) Check again to reassure if colnames are identical before binding
```{r}
#
identical(colnames(data.vertical$RPPAArray), colnames(data.vertical$RNASeq2GeneNorm))
identical(colnames(data.vertical$RNASeq2GeneNorm), colnames(data.vertical$miRNASeqGene))
```

## b) Merge data into 1 matrix
```{r}
data.vert.joined = do.call("rbind", list(data.vertical$RPPAArray, data.vertical$RNASeq2GeneNorm, data.vertical$miRNASeqGene))
```
```{r}
data.matrix = as.matrix(data.vert.joined)
```


# 1. Bicluster hierarchy

## A. Run 20 FABIA of 10 Biclusters and compare the results

```{r}
#Number of runs
n.runs = 20
```

```{r}
# Generate one seed per run
set.seed(42)
seeds = sample.int(100, n.runs)
seeds
```

```{r}
#Number of Biclusters
n.bcs = 10

results.FABIA = list()

for (i in 1:n.runs) {
  set.seed(seeds[i])
  fabia.curr = fabia(data.matrix, p =  n.bcs)
  results.FABIA[[i]] = BiclustSet(fabia.curr)
}
```

```{r}
result.FABIA.all = results.FABIA[[1]]

cols = result.FABIA.all@ColumnMembership
genes = result.FABIA.all@GenesMembership

for (i in 2:n.runs) {
  cols = rbind(cols, results.FABIA[[i]]@ColumnMembership)
  genes = cbind(genes, results.FABIA[[i]]@GenesMembership)
}
```

```{r}
result.FABIA.all@ColumnMembership = cols
result.FABIA.all@GenesMembership = genes
result.FABIA.all@Number = n.bcs * n.runs
#View(result.FABIA.all)
```

```{r, fig.width=20, fig.height=10}
jaccard.FABIA = similarity(result.FABIA.all, index = "jaccard", type = "both")
clustree.FABIA = HCLtree(jaccard.FABIA)
plot(clustree.FABIA, cex = 0.5)
abline(h = 0.9, col = 2, lwd = 0.5)
abline(h = 0.995, col = 3, lwd = 0.5)
```
```{r}
biclusters = cutree(clustree.FABIA,h=0.995)
nBiC = length(unique(biclusters))
resNames = list()
sampleNames = list()
```

```{r}
#Which nodes the 200 FABIA biclusters belong in
biclusters
```


```{r}
for(l in 1:nBiC){
  #print(l)
  idx = which(biclusters==l)
  #if (length(idx) == 1) {
  #  print(l)
  #}
  geneInfo = result.FABIA.all@GenesMembership[,idx, drop=F]
  sampleInfo = result.FABIA.all@ColumnMembership[idx,, drop=F]

  outGenes = unlist(apply(geneInfo, 1, function(x) length(which(x)==TRUE)))
  outSamples = unlist(apply(sampleInfo, 2, function(x) length(which(x)==TRUE)))
  #print(l)
  #print(outGenes)
  resNames[[l]] = rownames(data.matrix)[outGenes==length(idx)] # Originally a >= sign, but > is not possible.
  sampleNames[[l]] = colnames(data.matrix)[outSamples>=length(idx)/2]
}
```

```{r}
out.superbiclust.FABIA = list("resNames" = resNames,
                              "sampleNames" = sampleNames,
                              "biclust.per.node" = table(biclusters))
saveRDS(out.superbiclust.FABIA, file = "data/sel_features/results_superbiclust.fabia.rds")
```


## B. Compare FABIA and MOFA results

```{r}
#Load existing FABIA data
models.fabia = readRDS("data/out_models/results_fabia.rds")
fabia_10 =  models.fabia$`10bc`
result.FABIA = BiclustSet(fabia_10)
```

```{r}
#Load existing MOFA data
filepath = "data/out_models/mofa10f.hdf5"
MOFA_10 = load_model(filepath)
```

```{r}
#Extract MOFA outputs to that of the same format as FABIA

#factors.MOFA = get_factors(model.MOFA)
Z.MOFA = t(MOFA_10@expectations$Z$group1)
#Z.MOFA = get_factors(model.MOFA)$group1
L.MOFA.list = MOFA_10@expectations$W
L.MOFA = rbind(L.MOFA.list$view_RPPA, L.MOFA.list$view_R2Gn, L.MOFA.list$view_mRNA)
```

The Z and L threasholds are defined and calculated with the same analogy as for FABIA.

The code below is copied from the extractBic function in the FABIA package, which is used by superbiclust to convert FABIA results to bicluster form.

```{r}
threasZ.MOFA = 0.5

tt = 0
for (i in 1:ncol(L.MOFA)) {
  tt = tt + sum(L.MOFA[,i]^2) * sum(Z.MOFA[i,]^2)
}
tt = tt/(as.double(nrow(L.MOFA)) * as.double(ncol(L.MOFA)) * as.double(ncol(Z.MOFA)))
threasL.MOFA = sqrt(tt) / threasZ.MOFA
threasL.MOFA
```


For the sake of fairness, we will be calculating FABIA Z and L thresholds using the same manual way.

```{r}
threasZ.fabia= 0.5

tt = 0
for (i in 1:ncol(L(fabia_10))) {
  tt = tt + sum(L(fabia_10)[,i]^2) * sum(Z(fabia_10)[i,]^2)
}
tt = tt/(as.double(nrow(L(fabia_10))) * as.double(ncol(L(fabia_10))) * as.double(ncol(Z(fabia_10))))
threasL.fabia = sqrt(tt) / threasZ.fabia
threasL.fabia
```

```{r}
result.MOFA_FABIA_manual = result.FABIA
result.MOFA_FABIA_manual@ColumnMembership = rbind(((abs(Z(fabia_10)) > threasZ.fabia)), (abs(Z.MOFA) > threasZ.MOFA))
result.MOFA_FABIA_manual@GenesMembership = cbind(((abs(L(fabia_10)) > threasL.fabia)), (abs(L.MOFA) > threasL.MOFA))
result.MOFA_FABIA_manual@Number = 10*2
```

```{r}
jaccard.MOFA_FABIA_manual = similarity(result.MOFA_FABIA_manual, index = "jaccard", type = "both")
clustree.MOFA_FABIA_manual = HCLtree(jaccard.MOFA_FABIA_manual)
plot(clustree.MOFA_FABIA_manual)
abline(h = 0.9, col = 2, lwd = 0.5)
abline(h = 0.995, col = 3, lwd = 0.5)
```

### Important: FABIA biclusters are not in the same order as Superbiclust 1 ~ 10

The correct order is as follows:
```{r}
setNames(colnames(L(fabia_10)), c(1:10))
```


```{r}
biclusters.fm = cutree(clustree.MOFA_FABIA_manual,h=0.9)
nBiC.fm = length(unique(biclusters.fm))
resNames.fm = list()
sampleNames.fm = list()
```

```{r}
#Which nodes the 10 FABIA biclusters and 10 MOFA factors belong in
biclusters.fm
```


```{r}
for(l in 1:nBiC.fm){

  idx = which(biclusters.fm==l)

  geneInfo = result.MOFA_FABIA_manual@GenesMembership[,idx, drop=F]
  sampleInfo = result.MOFA_FABIA_manual@ColumnMembership[idx,, drop=F]

  outGenes = unlist(apply(geneInfo, 1, function(x) length(which(x)==TRUE)))
  outSamples = unlist(apply(sampleInfo, 2, function(x) length(which(x)==TRUE)))
  resNames.fm[[l]] = rownames(data.matrix)[outGenes==length(idx)]
  sampleNames.fm[[l]] = colnames(data.matrix)[outSamples>=length(idx)/2]
}
```

```{r}
out.superbiclust.FABIA_MOFA = list("resNames" = resNames.fm,
                              "sampleNames" = sampleNames.fm,
                              "biclust.per.node" = table(biclusters.fm))
saveRDS(out.superbiclust.FABIA_MOFA, file = "data/sel_features/results_superbiclust.fabia_mofa.rds")
```
