---
title: "Gaussian Graphical Model Lambda Training"
author: "John Zhuang"
date: "2024-08-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rags2ridges)
```

The purpose of this document is to separate optimal lambda exploration from the main GGM training document,

so that rendering time could be saved for the main document.

# 0. Data Preparation

```{r}
data.horizontal = readRDS("data/derived/data_horizontal_joined.rds")
data.numeric = as.matrix(data.horizontal[, 1:701])
data.Y = data.horizontal$vital_status
```

```{r}
group.mRSG = 1:470
group.R2Gn = 471:668
group.RPPA = 669:701
```

# 1. Wholistic analysis by omic 

## A. RPPAArray

```{r}
data.RPPA = data.numeric[, group.RPPA]
```

```{r}
set.seed(42)
opt.RPPA = optPenalty.kCVauto(Y = data.RPPA, lambdaMin = 1e-11, lambdaMax = 10)
opt.RPPA.10 = optPenalty.kCVauto(Y = data.RPPA, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.RPPA.5 = optPenalty.kCVauto(Y = data.RPPA, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.RPPA.3 = optPenalty.kCVauto(Y = data.RPPA, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

Lambdas.RPPA = setNames(c(opt.RPPA.3$optLambda, opt.RPPA.5$optLambda, opt.RPPA.10$optLambda, opt.RPPA$optLambda), c(3,5,10,43))
Lambdas.RPPA
```
## B. RNASeq2GeneNorm

```{r}
data.R2Gn = data.numeric[, group.R2Gn]

set.seed(42)
opt.R2Gn = optPenalty.kCVauto(Y = data.R2Gn, lambdaMin = 1e-11, lambdaMax = 10)
opt.R2Gn.10 = optPenalty.kCVauto(Y = data.R2Gn, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.R2Gn.5 = optPenalty.kCVauto(Y = data.R2Gn, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.R2Gn.3 = optPenalty.kCVauto(Y = data.R2Gn, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

Lambdas.R2Gn = setNames(c(opt.R2Gn.3$optLambda, opt.R2Gn.5$optLambda, opt.R2Gn.10$optLambda, opt.R2Gn$optLambda), c(3,5,10,43))
Lambdas.R2Gn
```
## C. miRNASeqGene

```{r}
data.mRSG = data.numeric[, group.mRSG]

set.seed(42)
opt.mRSG = optPenalty.kCVauto(Y = data.mRSG, lambdaMin = 1e-11, lambdaMax = 10)
opt.mRSG.10 = optPenalty.kCVauto(Y = data.mRSG, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.mRSG.5 = optPenalty.kCVauto(Y = data.mRSG, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.mRSG.3 = optPenalty.kCVauto(Y = data.mRSG, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

Lambdas.mRSG = setNames(c(opt.mRSG.3$optLambda, opt.mRSG.5$optLambda, opt.mRSG.10$optLambda, opt.mRSG$optLambda), c(3,5,10,43))
Lambdas.mRSG
```

## D. All data

```{r}
set.seed(42)
opt.all = optPenalty.kCVauto(Y = data.numeric, lambdaMin = 1e-11, lambdaMax = 10)
opt.all.10 = optPenalty.kCVauto(Y = data.numeric, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.all.5 = optPenalty.kCVauto(Y = data.numeric, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.all.3 = optPenalty.kCVauto(Y = data.numeric, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

Lambdas.all = setNames(c(opt.all.3$optLambda, opt.all.5$optLambda, opt.all.10$optLambda, opt.all$optLambda), c(3,5,10,43))
Lambdas.all
```


# 2. Split data into alive and dead
## A. RPPAArray

```{r}
data.RPPA.alive = data.RPPA[which(data.Y == 0), ]
data.RPPA.dead = data.RPPA[which(data.Y == 1), ]
```

```{r}
set.seed(42)
opt.RPPA.alive = optPenalty.kCVauto(Y = data.RPPA.alive, lambdaMin = 1e-11, lambdaMax = 10)
opt.RPPA.alive.10 = optPenalty.kCVauto(Y = data.RPPA.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.RPPA.alive.5 = optPenalty.kCVauto(Y = data.RPPA.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.RPPA.alive.3 = optPenalty.kCVauto(Y = data.RPPA.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

opt.RPPA.dead = optPenalty.kCVauto(Y = data.RPPA.dead, lambdaMin = 1e-11, lambdaMax = 10)
opt.RPPA.dead.10 = optPenalty.kCVauto(Y = data.RPPA.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.RPPA.dead.5 = optPenalty.kCVauto(Y = data.RPPA.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.RPPA.dead.3 = optPenalty.kCVauto(Y = data.RPPA.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

Lambdas.RPPA.alive = setNames(c(opt.RPPA.alive.3$optLambda, opt.RPPA.alive.5$optLambda, opt.RPPA.alive.10$optLambda, opt.RPPA.alive$optLambda), c(3,5,10,43))
Lambdas.RPPA.dead = setNames(c(opt.RPPA.dead.3$optLambda, opt.RPPA.dead.5$optLambda, opt.RPPA.dead.10$optLambda, opt.RPPA.dead$optLambda), c(3,5,10,43))

Lambdas.RPPA.alive
Lambdas.RPPA.dead
```

## B. RNASeq2GeneNorm


```{r}
data.R2Gn.alive = data.R2Gn[which(data.Y == 0), ]
data.R2Gn.dead = data.R2Gn[which(data.Y == 1), ]
```

```{r}
set.seed(42)
opt.R2Gn.alive = optPenalty.kCVauto(Y = data.R2Gn.alive, lambdaMin = 1e-11, lambdaMax = 10)
opt.R2Gn.alive.10 = optPenalty.kCVauto(Y = data.R2Gn.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.R2Gn.alive.5 = optPenalty.kCVauto(Y = data.R2Gn.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.R2Gn.alive.3 = optPenalty.kCVauto(Y = data.R2Gn.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

opt.R2Gn.dead = optPenalty.kCVauto(Y = data.R2Gn.dead, lambdaMin = 1e-11, lambdaMax = 10)
opt.R2Gn.dead.10 = optPenalty.kCVauto(Y = data.R2Gn.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.R2Gn.dead.5 = optPenalty.kCVauto(Y = data.R2Gn.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.R2Gn.dead.3 = optPenalty.kCVauto(Y = data.R2Gn.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

Lambdas.R2Gn.alive = setNames(c(opt.R2Gn.alive.3$optLambda, opt.R2Gn.alive.5$optLambda, opt.R2Gn.alive.10$optLambda, opt.R2Gn.alive$optLambda), c(3,5,10,43))
Lambdas.R2Gn.dead = setNames(c(opt.R2Gn.dead.3$optLambda, opt.R2Gn.dead.5$optLambda, opt.R2Gn.dead.10$optLambda, opt.R2Gn.dead$optLambda), c(3,5,10,43))

Lambdas.R2Gn.alive
Lambdas.R2Gn.dead
```

## C. RPPA Array + RNASeq2GeneNorm

```{r}
data.PR = data.numeric[, c(group.RPPA, group.R2Gn)]

data.PR.alive = data.PR[which(data.Y == 0), ]
data.PR.dead = data.PR[which(data.Y == 1), ]
```

```{r}
set.seed(42)
opt.PR.alive = optPenalty.kCVauto(Y = data.PR.alive, lambdaMin = 1e-11, lambdaMax = 10)
opt.PR.alive.10 = optPenalty.kCVauto(Y = data.PR.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.PR.alive.5 = optPenalty.kCVauto(Y = data.PR.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.PR.alive.3 = optPenalty.kCVauto(Y = data.PR.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

opt.PR.dead = optPenalty.kCVauto(Y = data.PR.dead, lambdaMin = 1e-11, lambdaMax = 10)
opt.PR.dead.10 = optPenalty.kCVauto(Y = data.PR.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.PR.dead.5 = optPenalty.kCVauto(Y = data.PR.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.PR.dead.3 = optPenalty.kCVauto(Y = data.PR.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

Lambdas.PR.alive = setNames(c(opt.PR.alive.3$optLambda, opt.PR.alive.5$optLambda, opt.PR.alive.10$optLambda, opt.PR.alive$optLambda), c(3,5,10,43))
Lambdas.PR.dead = setNames(c(opt.PR.dead.3$optLambda, opt.PR.dead.5$optLambda, opt.PR.dead.10$optLambda, opt.PR.dead$optLambda), c(3,5,10,43))

Lambdas.PR.alive
Lambdas.PR.dead
```

## D. miRNASeqGene

```{r}
data.mRSG.alive = data.mRSG[which(data.Y == 0), ]
data.mRSG.dead = data.mRSG[which(data.Y == 1), ]
```

```{r}
set.seed(42)
opt.mRSG.alive = optPenalty.kCVauto(Y = data.mRSG.alive, lambdaMin = 1e-11, lambdaMax = 10)
opt.mRSG.alive.10 = optPenalty.kCVauto(Y = data.mRSG.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.mRSG.alive.5 = optPenalty.kCVauto(Y = data.mRSG.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.mRSG.alive.3 = optPenalty.kCVauto(Y = data.mRSG.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

opt.mRSG.dead = optPenalty.kCVauto(Y = data.mRSG.dead, lambdaMin = 1e-11, lambdaMax = 10)
opt.mRSG.dead.10 = optPenalty.kCVauto(Y = data.mRSG.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.mRSG.dead.5 = optPenalty.kCVauto(Y = data.mRSG.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.mRSG.dead.3 = optPenalty.kCVauto(Y = data.mRSG.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

Lambdas.mRSG.alive = setNames(c(opt.mRSG.alive.3$optLambda, opt.mRSG.alive.5$optLambda, opt.mRSG.alive.10$optLambda, opt.mRSG.alive$optLambda), c(3,5,10,43))
Lambdas.mRSG.dead = setNames(c(opt.mRSG.dead.3$optLambda, opt.mRSG.dead.5$optLambda, opt.mRSG.dead.10$optLambda, opt.mRSG.dead$optLambda), c(3,5,10,43))

Lambdas.mRSG.alive
Lambdas.mRSG.dead
```

## E. All Data

```{r}
data.all.alive = data.numeric[which(data.Y == 0), ]
data.all.dead = data.numeric[which(data.Y == 1), ]
```

```{r}
set.seed(42)
opt.all.alive = optPenalty.kCVauto(Y = data.all.alive, lambdaMin = 1e-11, lambdaMax = 10)
opt.all.alive.10 = optPenalty.kCVauto(Y = data.all.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.all.alive.5 = optPenalty.kCVauto(Y = data.all.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.all.alive.3 = optPenalty.kCVauto(Y = data.all.alive, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

opt.all.dead = optPenalty.kCVauto(Y = data.all.dead, lambdaMin = 1e-11, lambdaMax = 10)
opt.all.dead.10 = optPenalty.kCVauto(Y = data.all.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 10)
opt.all.dead.5 = optPenalty.kCVauto(Y = data.all.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 5)
opt.all.dead.3 = optPenalty.kCVauto(Y = data.all.dead, lambdaMin = 1e-11, lambdaMax = 10, fold = 3)

Lambdas.all.alive = setNames(c(opt.all.alive.3$optLambda, opt.all.alive.5$optLambda, opt.all.alive.10$optLambda, opt.all.alive$optLambda), c(3,5,10,43))
Lambdas.all.dead = setNames(c(opt.all.dead.3$optLambda, opt.all.dead.5$optLambda, opt.all.dead.10$optLambda, opt.all.dead$optLambda), c(3,5,10,43))

Lambdas.all.alive
Lambdas.all.dead
```

# 3. Make a dataframe of all optLambda values and export it

```{r}
ggm.optLambdas = data.frame(c(3,5,10,43),
                            Lambdas.RPPA,
                            Lambdas.R2Gn,
                            Lambdas.mRSG,
                            Lambdas.all,
                            Lambdas.RPPA.alive,
                            Lambdas.RPPA.dead,
                            Lambdas.R2Gn.alive,
                            Lambdas.R2Gn.dead,
                            Lambdas.PR.alive,
                            Lambdas.PR.dead,
                            Lambdas.mRSG.alive,
                            Lambdas.mRSG.dead,
                            Lambdas.all.alive,
                            Lambdas.all.dead)
ggm.optLambdas
```

```{r}
saveRDS(ggm.optLambdas, file = "data/temp/ggm_all_optLambdas.rds")
```

